---
title: "S&P 500 ESG Data Pipeline and Crash-Risk Warning System"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
params:
  start_year: 2012        # First year of the panel window
  end_year: 2023          # Last year of feature data; also serves as the external test feature year
  run_live_edgar_fill: false  # Set to true to download missing 10-K filings from SEC EDGAR (slow)
  fill_start_year: 2021   # Earliest year to backfill when live EDGAR is enabled
  fill_end_year: 2023     # Latest year to backfill when live EDGAR is enabled
  live_fill_max_companies: 10  # Cap on companies per live EDGAR run (0 = no limit)
  sec_user_agent: "Academic ESG Project research@example.com"  # Required by SEC for rate-limiting
---

## Goal

This notebook implements a complete S&P 500 ESG research pipeline with three integrated layers:

1. **Data Foundation**: Build a multi-year firm-year panel by merging the current S&P 500 constituent universe, daily price history (Yahoo Finance), a cross-sectional ESG snapshot, and annual EDGAR 10-K text-based ESG signals.
2. **Data Usability Audit**: Measure coverage, missingness, and sample usability under dual-denominator definitions (full 503-firm sample vs. effective post-index-entry sample), broken down by year and sector.
3. **Crash-Risk Warning System**: Train and validate supervised learning models (weighted logistic regression, probability random forest) across nested feature sets to predict one-year and two-year-ahead stock price crash events, with threshold-optimized company-level warning lists.

The core research question: **Can ESG signals — both snapshot ratings and text-derived keyword intensity from SEC filings — improve crash-risk prediction beyond price-based features alone?**

Important caveats:

- The panel is built from the **current** S&P 500 constituent set, so historical rows carry **survivorship bias**.
- The snapshot ESG source is cross-sectional (one observation per firm), not year-specific.
- EDGAR ESG scores are text-derived keyword intensity measures, not externally validated ratings.
- Model outputs should be treated as **screening signals**, not investment or credit decisions.

## Dependencies and Paths

```{r setup}
# ---------------------------------------------------------------------------
# Package installation and loading
# ---------------------------------------------------------------------------
# All packages are declared upfront so missing ones are auto-installed.
# This keeps the notebook self-contained on any machine with R and rmarkdown.
required_packages <- c(
  "dplyr",         # Data manipulation
  "readr",         # CSV reading/writing
  "stringr",       # String processing
  "purrr",         # Functional programming tools
  "tidyr",         # Data reshaping
  "tibble",        # Modern data frames
  "lubridate",     # Date handling
  "jsonlite",      # JSON parsing (SEC EDGAR API)
  "xml2",          # XML/HTML parsing
  "rvest",         # Web scraping (Wikipedia)
  "curl",          # HTTP requests
  "janitor",       # Data cleaning helpers
  "knitr",         # Notebook rendering
  "rmarkdown",     # R Markdown engine
  "data.table",    # Fast CSV I/O via fread/fwrite
  "ggplot2",       # Visualization
  "scales",        # Axis formatting
  "ranger",        # Random forest (probability estimation)
  "tidyquant",     # Financial data utilities
  "quantmod",      # Quantitative finance tools
  "zoo"            # Time series utilities
)

missing_packages <- setdiff(required_packages, rownames(installed.packages()))

if (length(missing_packages) > 0) {
  install.packages(missing_packages, repos = "https://cloud.r-project.org/")
}

knitr::opts_chunk$set(
  echo = TRUE,
  warning = TRUE,
  message = TRUE
)

options(stringsAsFactors = FALSE)

suppressPackageStartupMessages({
  library(dplyr)
  library(readr)
  library(stringr)
  library(purrr)
  library(tidyr)
  library(tibble)
  library(lubridate)
  library(jsonlite)
  library(xml2)
  library(rvest)
  library(curl)
  library(janitor)
  library(data.table)
  library(ggplot2)
  library(scales)
  library(zoo)
})

# ---------------------------------------------------------------------------
# Project path resolution
# ---------------------------------------------------------------------------
# The notebook auto-detects its own directory so relative paths work
# regardless of where knitr is invoked from.

get_document_dir <- function() {
  input_path <- tryCatch(knitr::current_input(dir = TRUE), error = function(error) "")

  if (nzchar(input_path) && file.exists(input_path)) {
    return(dirname(normalizePath(input_path, winslash = "/", mustWork = TRUE)))
  }

  normalizePath(getwd(), winslash = "/", mustWork = TRUE)
}

PROJECT_ROOT <- get_document_dir()
DATA_DIR <- file.path(PROJECT_ROOT, "data")
RAW_DIR <- file.path(DATA_DIR, "raw")
DERIVED_DIR <- file.path(DATA_DIR, "derived")
OUTPUT_DIR <- file.path(PROJECT_ROOT, "output")
OUTPUT_USABILITY_DIR <- file.path(OUTPUT_DIR, "01_data_usability")
OUTPUT_ML_DIR <- file.path(OUTPUT_DIR, "02_ml_warning")

# Ensure all output directories exist before any writes
dir.create(DATA_DIR, recursive = TRUE, showWarnings = FALSE)
dir.create(RAW_DIR, recursive = TRUE, showWarnings = FALSE)
dir.create(DERIVED_DIR, recursive = TRUE, showWarnings = FALSE)
dir.create(OUTPUT_DIR, recursive = TRUE, showWarnings = FALSE)
dir.create(OUTPUT_USABILITY_DIR, recursive = TRUE, showWarnings = FALSE)
dir.create(OUTPUT_ML_DIR, recursive = TRUE, showWarnings = FALSE)

# ---------------------------------------------------------------------------
# Global parameters — derived from YAML header or environment variables
# ---------------------------------------------------------------------------
START_YEAR <- as.integer(params$start_year)                   # 2012
END_YEAR <- as.integer(params$end_year)                       # 2023 — last feature year
WARNING_EXTERNAL_START_YEAR <- 2024L                          # First year of external test window
WARNING_EXTERNAL_END_YEAR <- 2025L                            # Last year of external test window
WARNING_LOOKAHEAD_YEARS <- 2L                                 # Number of forward years for 2Y labels
# Environment flags allow forced recomputation of cached price/label files
FORCE_REFRESH_WARNING_PRICES <- identical(Sys.getenv("FORCE_REFRESH_WARNING_PRICES"), "1") ||
  identical(Sys.getenv("FORCE_REFRESH_2024_PRICES"), "1") ||
  identical(Sys.getenv("FORCE_REFRESH_2025_PRICES"), "1")
FORCE_REBUILD_WARNING_LABELS <- identical(Sys.getenv("FORCE_REBUILD_WARNING_LABELS"), "1")

# ---------------------------------------------------------------------------
# External data source URLs and headers
# ---------------------------------------------------------------------------
WIKI_URL <- "https://en.wikipedia.org/wiki/List_of_S%26P_500_companies"
# SEC requires a User-Agent header identifying the requester
SEC_HEADERS <- c(
  "User-Agent" = params$sec_user_agent,
  "Accept" = "application/json, text/html;q=0.9"
)

# All required source files live under data/raw so the project is portable.
# If any are missing, the notebook halts immediately with a clear error.
EXTERNAL_SOURCES <- list(
  snapshot_esg = file.path(RAW_DIR, "SP 500 ESG Risk Ratings.csv"),
  universe_cache = file.path(RAW_DIR, "company_universe.csv"),
  price_cache = file.path(RAW_DIR, "prices_daily.csv"),
  edgar_cache = file.path(RAW_DIR, "esg_signals_panel.csv")
)

missing_source_files <- unname(EXTERNAL_SOURCES[!vapply(EXTERNAL_SOURCES, file.exists, logical(1))])

if (length(missing_source_files) > 0) {
  stop(sprintf(
    "Missing required local source file(s) under data/raw: %s",
    paste(basename(missing_source_files), collapse = ", ")
  ))
}

# ---------------------------------------------------------------------------
# Output file path definitions — all derived and output files are named here
# so the notebook has a single source of truth for every write/read path.
# ---------------------------------------------------------------------------

# Stage 1: Data foundation outputs
OUTPUT_FILES <- list(
  constituents = file.path(DERIVED_DIR, "sp500_constituents_current.csv"),
  constituents_cache = file.path(DERIVED_DIR, "sp500_wiki_current.csv"),
  price_daily = file.path(DERIVED_DIR, "sp500_price_daily_2011_2023.csv"),
  price_annual = file.path(DERIVED_DIR, "sp500_price_annual_features_2012_2023.csv"),
  edgar_yearly = file.path(DERIVED_DIR, "sp500_edgar_esg_annual_2012_2023.csv"),
  panel = file.path(DERIVED_DIR, "sp500_price_esg_panel_2012_2023.csv"),
  coverage_year = file.path(OUTPUT_DIR, "sp500_coverage_by_year.csv"),
  coverage_overview = file.path(OUTPUT_DIR, "sp500_coverage_overview.csv"),
  coverage_company = file.path(OUTPUT_DIR, "sp500_company_coverage_flags.csv")
)

# Stage 2: Usability audit outputs (tables + charts)
OUTPUT_USABILITY_FILES <- list(
  overview = file.path(OUTPUT_USABILITY_DIR, "01_overview_dual_denominator.csv"),
  coverage_year = file.path(OUTPUT_USABILITY_DIR, "02_coverage_by_year_dual_denominator.csv"),
  coverage_sector = file.path(OUTPUT_USABILITY_DIR, "03_coverage_by_sector_dual_denominator.csv"),
  field_missingness = file.path(OUTPUT_USABILITY_DIR, "04_field_missingness.csv"),
  company_buckets = file.path(OUTPUT_USABILITY_DIR, "05_company_coverage_buckets.csv"),
  special_cases = file.path(OUTPUT_USABILITY_DIR, "06_special_cases_summary.csv"),
  yearly_lines = file.path(OUTPUT_USABILITY_DIR, "11_yearly_coverage_lines.png"),
  joint_bars = file.path(OUTPUT_USABILITY_DIR, "12_yearly_joint_coverage_bars.png"),
  sector_heatmap = file.path(OUTPUT_USABILITY_DIR, "13_sector_coverage_heatmap.png"),
  missingness_bar = file.path(OUTPUT_USABILITY_DIR, "14_field_missingness_bar.png"),
  bucket_bar = file.path(OUTPUT_USABILITY_DIR, "15_company_bucket_bar.png")
)

# Stage 3A: Derived ML files (price extensions, crash labels, modeling panels)
DERIVED_ML_FILES <- list(
  price_2024 = file.path(DERIVED_DIR, "sp500_price_daily_2024.csv"),
  price_2025 = file.path(DERIVED_DIR, "sp500_price_daily_2025.csv"),
  price_daily_extended = file.path(DERIVED_DIR, "sp500_price_daily_2011_2025.csv"),
  crash_metrics_extended = file.path(DERIVED_DIR, "sp500_crash_metrics_2012_2025.csv"),
  ml_panel = file.path(DERIVED_DIR, "sp500_ml_panel_2y_2012_2023_features_to_2024_2025_labels.csv")
)

# Stage 3B: ML output files (metrics, predictions, charts)
OUTPUT_ML_FILES <- list(
  training_panel = file.path(OUTPUT_ML_DIR, "01_training_panel_2y.csv"),
  validation_fold_metrics = file.path(OUTPUT_ML_DIR, "02_validation_fold_metrics_2y.csv"),
  threshold_tuning = file.path(OUTPUT_ML_DIR, "03_threshold_tuning_2y.csv"),
  cv_model_comparison = file.path(OUTPUT_ML_DIR, "04_model_comparison_cv_2y.csv"),
  test_metrics = file.path(OUTPUT_ML_DIR, "05_external_test_metrics_2024_2025_2y.csv"),
  confusion_matrix = file.path(OUTPUT_ML_DIR, "06_confusion_matrix_by_model_2y.csv"),
  company_predictions = file.path(OUTPUT_ML_DIR, "07_company_predictions_2024_2025_2y.csv"),
  top_risk_companies = file.path(OUTPUT_ML_DIR, "08_high_risk_companies_2024_2025_2y.csv"),
  pr_curve = file.path(OUTPUT_ML_DIR, "11_pr_curve_2y.png"),
  accuracy_precision_plot = file.path(OUTPUT_ML_DIR, "12_accuracy_precision_comparison_2y.png"),
  confusion_all_plot = file.path(OUTPUT_ML_DIR, "13_confusion_matrix_all_models_2y.png"),
  confusion_champion_plot = file.path(OUTPUT_ML_DIR, "14_confusion_matrix_champion_2y.png"),
  top_risk_plot = file.path(OUTPUT_ML_DIR, "15_top_risk_companies_2y.png"),
  logistic_coef_plot = file.path(OUTPUT_ML_DIR, "16_logistic_coefficients_2y.png"),
  forest_importance_plot = file.path(OUTPUT_ML_DIR, "17_forest_importance_2y.png")
)

message("Project root detected.")
message(sprintf("Snapshot ESG source file: %s", basename(EXTERNAL_SOURCES$snapshot_esg)))
message(sprintf("Universe cache source file: %s", basename(EXTERNAL_SOURCES$universe_cache)))
message(sprintf("Price cache source file: %s", basename(EXTERNAL_SOURCES$price_cache)))
message(sprintf("EDGAR cache source file: %s", basename(EXTERNAL_SOURCES$edgar_cache)))
```
## Helper Functions

```{r helpers}
# ---------------------------------------------------------------------------
# Path and identifier utilities
# ---------------------------------------------------------------------------

# Strip the project root prefix from absolute paths for display purposes
to_project_relative <- function(paths) {
  normalized_paths <- normalizePath(paths, winslash = "/", mustWork = FALSE)
  project_prefix <- paste0(normalizePath(PROJECT_ROOT, winslash = "/", mustWork = TRUE), "/")
  stringr::str_replace(normalized_paths, stringr::fixed(project_prefix), "")
}

# Standardize ticker symbols: trim whitespace and replace dots with dashes.
# Yahoo Finance and Wikipedia sometimes differ on BRK.B vs BRK-B, etc.
normalize_symbol <- function(x) {
  x %>%
    stringr::str_trim() %>%
    stringr::str_replace_all("\\.", "-")
}

# Pad CIK (Central Index Key) to SEC's canonical 10-digit zero-padded format.
# CIK values appear with or without leading zeros across different sources.
format_cik <- function(x) {
  raw_cik <- stringr::str_extract(as.character(x), "\\d+")
  stringr::str_pad(raw_cik, width = 10, side = "left", pad = "0")
}

# ---------------------------------------------------------------------------
# HTTP/SEC EDGAR helpers
# ---------------------------------------------------------------------------

# Fetch raw binary content from a URL, optionally limiting the download size.
# Used for SEC filings where document bodies can be very large.
fetch_binary <- function(url, headers = NULL, max_bytes = NULL) {
  handle <- if (is.null(headers) || length(headers) == 0) {
    curl::new_handle()
  } else {
    curl::new_handle(httpheader = headers)
  }

  if (is.null(max_bytes)) {
    response <- curl::curl_fetch_memory(url, handle = handle)
    return(response$content)
  }

  connection <- curl::curl(url, handle = handle)
  on.exit(close(connection), add = TRUE)
  readBin(connection, what = "raw", n = max_bytes)
}

fetch_html <- function(url, headers = NULL) {
  xml2::read_html(rawToChar(fetch_binary(url, headers = headers)))
}

fetch_json <- function(url, headers = NULL) {
  jsonlite::fromJSON(rawToChar(fetch_binary(url, headers = headers)), simplifyVector = FALSE)
}

# ---------------------------------------------------------------------------
# Financial computation helpers
# ---------------------------------------------------------------------------

# Maximum drawdown: largest peak-to-trough decline in a price series.
# Returns a negative number (e.g., -0.35 means a 35% drawdown).
calc_max_drawdown <- function(prices) {
  prices <- as.numeric(prices)
  prices <- prices[is.finite(prices)]

  if (length(prices) < 2) {
    return(NA_real_)
  }

  drawdowns <- prices / cummax(prices) - 1
  min(drawdowns, na.rm = TRUE)
}

# ---------------------------------------------------------------------------
# EDGAR 10-K text processing and ESG keyword extraction
# ---------------------------------------------------------------------------

# Convert raw SEC filing HTML to clean, lowercased plain text.
# Strips scripts, styles, inline XBRL tags, and non-alphanumeric characters.
strip_html_text <- function(html_text) {
  clean_html <- html_text %>%
    stringr::str_replace_all("(?is)<script.*?</script>", " ") %>%
    stringr::str_replace_all("(?is)<style.*?</style>", " ") %>%
    stringr::str_replace_all("(?is)</?ix:[^>]+>", " ")

  plain_text <- tryCatch(
    xml2::xml_text(xml2::read_html(clean_html)),
    error = function(error) stringr::str_replace_all(clean_html, "<[^>]+>", " ")
  )

  plain_text %>%
    stringr::str_to_lower() %>%
    stringr::str_replace_all("[^a-z0-9\\s]", " ") %>%
    stringr::str_squish()
}

# Count total keyword occurrences in text (exact substring matching, case-insensitive).
# Each keyword is counted independently; overlapping hits are not double-counted
# since stringr::str_count with fixed() is non-overlapping by default.
count_keyword_hits <- function(text, keywords) {
  sum(purrr::map_int(
    keywords,
    ~ stringr::str_count(text, stringr::fixed(stringr::str_to_lower(.x)))
  ))
}

# ESG keyword dictionaries used for text-based signal extraction from 10-K filings.
# Keywords are domain-specific terms drawn from academic ESG disclosure research.
# The three dimensions follow standard ESG pillar definitions.
edgar_keywords <- list(
  e_keywords = c(
    "climate change",
    "greenhouse gas",
    "carbon emission",
    "carbon footprint",
    "renewable energy",
    "environmental sustainability",
    "pollution control",
    "net zero",
    "clean energy",
    "energy efficiency",
    "biodiversity",
    "water conservation",
    "waste reduction",
    "environmental impact",
    "climate risk",
    "carbon neutral",
    "scope 1",
    "scope 2",
    "scope 3"
  ),
  s_keywords = c(
    "employee wellbeing",
    "workplace safety",
    "diversity and inclusion",
    "human rights",
    "community investment",
    "labor practices",
    "gender diversity",
    "health and safety",
    "supply chain responsibility",
    "workforce development",
    "equal opportunity",
    "social responsibility",
    "pay equity",
    "employee benefits"
  ),
  g_keywords = c(
    "corporate governance",
    "board independence",
    "audit committee",
    "executive compensation",
    "whistleblower",
    "anti-corruption",
    "regulatory compliance",
    "shareholder rights",
    "board diversity",
    "risk oversight",
    "code of ethics",
    "independent director"
  )
)

# Get the SEC EDGAR submission index for a given CIK.
# Returns a table of 10-K and 10-K/A filings with accession numbers and dates.
# For firms with multiple filings in the same fiscal year, prefers 10-K over 10-K/A.
get_sec_submission_map <- function(cik, start_year, end_year, headers) {
  url <- sprintf("https://data.sec.gov/submissions/CIK%s.json", cik)
  payload <- tryCatch(fetch_json(url, headers = headers), error = function(error) NULL)

  if (is.null(payload)) {
    return(tibble())
  }

  recent <- payload$filings$recent
  if (is.null(recent$form)) {
    return(tibble())
  }

  recent_table <- tibble(
    form = unlist(recent$form),
    accession_number = unlist(recent$accessionNumber),
    filing_date = unlist(recent$filingDate),
    primary_document = unlist(recent$primaryDocument)
  ) %>%
    filter(form %in% c("10-K", "10-K/A")) %>%
    mutate(
      filing_date = as.Date(filing_date),
      filing_year = lubridate::year(filing_date),
      # Map filing date to the fiscal year it covers: filings Jan-Mar typically
      # cover the previous fiscal year (fiscal_year = filing_year - 1).
      fiscal_year = if_else(lubridate::month(filing_date) >= 4L, filing_year, filing_year - 1L),
      accession_nodash = stringr::str_replace_all(accession_number, "-", "")
    ) %>%
    filter(dplyr::between(fiscal_year, start_year, end_year)) %>%
    group_by(fiscal_year) %>%
    arrange(desc(form == "10-K/A"), .by_group = TRUE) %>%
    slice(1) %>%
    ungroup()

  recent_table
}

# Construct the direct URL to a specific SEC filing document.
# Format: https://www.sec.gov/Archives/edgar/data/{CIK-as-int}/{accession-nodash}/{primary-doc}
build_sec_filing_url <- function(cik, accession_nodash, primary_document) {
  sprintf(
    "https://www.sec.gov/Archives/edgar/data/%s/%s/%s",
    as.integer(cik),
    accession_nodash,
    primary_document
  )
}

# Download and parse one SEC 10-K filing to extract ESG keyword intensity scores.
# Scores are expressed as keyword hits per 1,000 filing words to normalize for
# document length. Returns NULL if the filing is too short (< 500 words) or
# unreachable, so the panel records these as missing EDGAR observations.
extract_one_edgar_row <- function(cik, symbol, fiscal_year, filing_info, headers) {
  filing_url <- build_sec_filing_url(
    cik = cik,
    accession_nodash = filing_info$accession_nodash,
    primary_document = filing_info$primary_document
  )

  # Cap download at ~800KB to avoid pulling massive exhibits
  filing_raw <- tryCatch(
    fetch_binary(filing_url, headers = headers, max_bytes = 800000L),
    error = function(error) NULL
  )

  if (is.null(filing_raw) || length(filing_raw) == 0) {
    return(NULL)
  }

  filing_text <- strip_html_text(rawToChar(filing_raw))
  word_count <- length(strsplit(filing_text, "\\s+")[[1]])

  if (!is.finite(word_count) || word_count < 500) {
    return(NULL)
  }

  # Count raw keyword hits for each ESG pillar, then scale per 1,000 words
  e_count <- count_keyword_hits(filing_text, edgar_keywords$e_keywords)
  s_count <- count_keyword_hits(filing_text, edgar_keywords$s_keywords)
  g_count <- count_keyword_hits(filing_text, edgar_keywords$g_keywords)
  scale <- max(word_count, 1) / 1000

  tibble(
    cik = cik,
    symbol = symbol,
    year = fiscal_year,
    e_score = round(e_count / scale, 4),
    s_score = round(s_count / scale, 4),
    g_score = round(g_count / scale, 4),
    esg_total = round((e_count + s_count + g_count) / scale, 4),
    e_count = e_count,
    s_count = s_count,
    g_count = g_count,
    word_count = word_count,
    filing_date = as.Date(filing_info$filing_date)
  )
}
```

## Load the Current S&P 500 Universe

This section builds the company master table by merging three sources:
1. A cross-sectional ESG snapshot (Risk Ratings CSV)
2. The current S&P 500 constituent list from Wikipedia
3. A local company universe cache for ticker-to-CIK mapping

All three are joined on normalized ticker symbols. When a field exists in multiple
sources, `coalesce` picks the first non-missing value, with Wikipedia preferred
for identifiers and the snapshot preferred for ESG metrics.

```{r sp500-universe}
# ---- Load and standardize the snapshot ESG ratings file ----
snapshot_source_full <- readr::read_csv(
  EXTERNAL_SOURCES$snapshot_esg,
  show_col_types = FALSE
)

# ---- Load the local company universe cache for ticker-to-CIK mapping ----
universe_map <- data.table::fread(
  EXTERNAL_SOURCES$universe_cache,
  showProgress = FALSE
) %>%
  as_tibble() %>%
  transmute(
    symbol = normalize_symbol(ticker),
    cik_universe = format_cik(cik),
    universe_company_name = company_name
  )

# ---- Load S&P 500 constituents: prefer cached Wikipedia scrape, re-fetch if missing ----
sp500_table <- if (file.exists(OUTPUT_FILES$constituents_cache)) {
  readr::read_csv(OUTPUT_FILES$constituents_cache, show_col_types = FALSE)
} else {
  tryCatch(
    fetch_html(WIKI_URL, headers = SEC_HEADERS) %>%
      rvest::html_element("table") %>%
      rvest::html_table(),
    error = function(error) NULL
  )
}

# ---- Standardize Wikipedia constituent fields ----
wiki_constituents <- if (!is.null(sp500_table)) {
  sp500_table %>%
    janitor::clean_names() %>%
    transmute(
      symbol_raw_wiki = symbol,
      symbol = normalize_symbol(symbol_raw_wiki),
      company_name_wiki = security,
      gics_sector_wiki = gics_sector,
      gics_sub_industry_wiki = gics_sub_industry,
      headquarters_location_wiki = headquarters_location,
      date_added = suppressWarnings(as.Date(date_added)),
      cik_wiki = format_cik(cik),
      founded = founded
    )
} else {
  tibble(
    symbol_raw_wiki = character(),
    symbol = character(),
    company_name_wiki = character(),
    gics_sector_wiki = character(),
    gics_sub_industry_wiki = character(),
    headquarters_location_wiki = character(),
    date_added = as.Date(character()),
    cik_wiki = character(),
    founded = character()
  )
}

# ---- Standardize snapshot ESG fields ----
snapshot_esg <- snapshot_source_full %>%
  janitor::clean_names() %>%
  transmute(
    symbol_raw_snapshot = symbol,
    symbol = normalize_symbol(symbol),
    snapshot_name = name,
    snapshot_sector = sector,
    snapshot_industry = industry,
    snapshot_headquarters = address,
    snapshot_total_esg_risk = total_esg_risk_score,
    snapshot_environment_risk = environment_risk_score,
    snapshot_governance_risk = governance_risk_score,
    snapshot_social_risk = social_risk_score,
    snapshot_controversy_level = controversy_level,
    snapshot_controversy_score = controversy_score,
    snapshot_esg_percentile = esg_risk_percentile,
    snapshot_esg_level = esg_risk_level
  )

# ---- Merge three sources on normalized symbol to build the company master table ----
# Priority order (via coalesce): Wikipedia > Snapshot > Universe cache
company_master <- snapshot_esg %>%
  left_join(wiki_constituents, by = "symbol") %>%
  left_join(universe_map, by = "symbol") %>%
  mutate(
    symbol_raw = coalesce(symbol_raw_wiki, symbol_raw_snapshot),
    company_name = coalesce(snapshot_name, company_name_wiki, universe_company_name),
    cik = coalesce(cik_wiki, cik_universe),
    gics_sector = coalesce(gics_sector_wiki, snapshot_sector),
    gics_sub_industry = coalesce(gics_sub_industry_wiki, snapshot_industry),
    headquarters_location = coalesce(headquarters_location_wiki, snapshot_headquarters),
    price_symbol = symbol,
    snapshot_available = !is.na(snapshot_total_esg_risk)
  ) %>%
  select(
    symbol,
    symbol_raw,
    company_name,
    cik,
    gics_sector,
    gics_sub_industry,
    headquarters_location,
    date_added,
    founded,
    snapshot_name,
    snapshot_sector,
    snapshot_industry,
    snapshot_total_esg_risk,
    snapshot_environment_risk,
    snapshot_governance_risk,
    snapshot_social_risk,
    snapshot_controversy_level,
    snapshot_controversy_score,
    snapshot_esg_percentile,
    snapshot_esg_level,
    price_symbol,
    snapshot_available
  )

data.table::fwrite(company_master, OUTPUT_FILES$constituents)

company_master %>%1
  summarise(
    constituent_rows = n(),
    unique_cik = n_distinct(cik),
    snapshot_esg_available = sum(snapshot_available, na.rm = TRUE)
  ) %>%
  knitr::kable()
```

## Load the Multi-Year Price Cache and Build Annual Features

This section reads the cached daily price history (Yahoo Finance), filters to
the current S&P 500 universe, and aggregates into annual features:
- **Return**: simple annual return (end_price / start_price - 1) and log return
- **Risk**: annualized volatility (daily log return SD * sqrt(252)) and max drawdown
- **Liquidity**: average and median daily volume
- **Trading intensity**: number of trading days observed

```{r prices}
# ---- Build a price-symbol lookup keyed to normalized ticker ----
price_map <- company_master %>%
  distinct(symbol, price_symbol, company_name, cik, gics_sector, gics_sub_industry)

# ---- Load the full price cache and filter to S&P 500 symbols ----
price_daily <- data.table::fread(
  EXTERNAL_SOURCES$price_cache,
  select = c("ticker", "date", "adj_close", "volume"),
  showProgress = FALSE
) %>%
  as_tibble() %>%
  transmute(
    price_symbol = normalize_symbol(ticker),
    date = as.Date(date),
    adj_close = as.numeric(adj_close),
    volume = as.numeric(volume)
  ) %>%
  filter(price_symbol %in% price_map$price_symbol) %>%
  inner_join(price_map, by = "price_symbol") %>%
  arrange(symbol, date)

price_daily %>%
  summarise(
    firms_with_prices = n_distinct(symbol),
    daily_rows = n(),
    first_date = min(date, na.rm = TRUE),
    last_date = max(date, na.rm = TRUE)
  ) %>%
  knitr::kable()

data.table::fwrite(price_daily, OUTPUT_FILES$price_daily)

# ---- Aggregate daily prices into annual features per company-year ----
annual_price_features <- price_daily %>%
  group_by(symbol) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(
    daily_return = adj_close / lag(adj_close) - 1,
    daily_log_return = log(adj_close / lag(adj_close)),
    year = lubridate::year(date)
  ) %>%
  ungroup() %>%
  filter(dplyr::between(year, START_YEAR, END_YEAR)) %>%
  group_by(symbol, company_name, cik, gics_sector, gics_sub_industry, year) %>%
  summarise(
    n_trading_days = n(),
    first_trade = min(date),
    last_trade = max(date),
    start_price = first(adj_close),
    end_price = last(adj_close),
    annual_return = end_price / start_price - 1,
    annual_log_return = sum(daily_log_return, na.rm = TRUE),
    annual_volatility = sd(daily_log_return, na.rm = TRUE) * sqrt(252),
    avg_volume = mean(volume, na.rm = TRUE),
    median_volume = median(volume, na.rm = TRUE),
    max_drawdown = calc_max_drawdown(adj_close),
    .groups = "drop"
  )

data.table::fwrite(annual_price_features, OUTPUT_FILES$price_annual)

annual_price_features %>%
  summarise(
    annual_rows = n(),
    firms_covered = n_distinct(symbol),
    years_covered = n_distinct(year)
  ) %>%
  knitr::kable()
```

## Load the Annual EDGAR ESG Cache

The EDGAR ESG panel provides annual keyword-intensity scores from 10-K filings.
Each row is one company-year with scores measured as keyword hits per 1,000 words.
The cache is pre-built; live SEC backfill is disabled by default (see next section).

```{r edgar-cache}
# ---- Load cached EDGAR ESG signals, filter to panel years ----
edgar_cached <- data.table::fread(
  EXTERNAL_SOURCES$edgar_cache,
  showProgress = FALSE
) %>%
  as_tibble() %>%
  transmute(
    cik = format_cik(cik),
    cached_symbol = normalize_symbol(ticker),
    year = as.integer(year),
    e_score = as.numeric(e_score),
    s_score = as.numeric(s_score),
    g_score = as.numeric(g_score),
    esg_total = as.numeric(esg_total),
    e_count = as.numeric(e_count),
    s_count = as.numeric(s_count),
    g_count = as.numeric(g_count),
    word_count = as.numeric(word_count),
    filing_date = as.Date(filing_date)
  ) %>%
  filter(dplyr::between(year, START_YEAR, END_YEAR)) %>%
  semi_join(company_master %>% distinct(cik), by = "cik") %>%
  arrange(cik, year) %>%
  distinct(cik, year, .keep_all = TRUE)

edgar_cached %>%
  summarise(
    edgar_rows = n(),
    firms_with_edgar = n_distinct(cik),
    first_year = min(year, na.rm = TRUE),
    last_year = max(year, na.rm = TRUE)
  ) %>%
  knitr::kable()
```

## Optional: Live Backfill of Missing EDGAR Records

This step is disabled by default because pulling 10-K text in bulk for roughly 500 firms is slow.

If you want to continue filling missing observations later, set `run_live_edgar_fill` to `true` in the YAML parameters and replace `sec_user_agent` with your own contact email.

```{r edgar-live-fill, eval = params$run_live_edgar_fill}
fill_years <- seq.int(as.integer(params$fill_start_year), as.integer(params$fill_end_year))

missing_company_years <- company_master %>%
  distinct(symbol, cik) %>%
  tidyr::crossing(year = fill_years) %>%
  anti_join(edgar_cached %>% distinct(cik, year), by = c("cik", "year")) %>%
  filter(!is.na(cik))

candidate_companies <- missing_company_years %>%
  distinct(symbol, cik) %>%
  {
    if (isTRUE(params$live_fill_max_companies > 0)) slice_head(., n = params$live_fill_max_companies) else .
  }

message(sprintf("Live EDGAR fill candidate companies: %d", nrow(candidate_companies)))

live_rows <- purrr::map_dfr(seq_len(nrow(candidate_companies)), function(index) {
  one_company <- candidate_companies[index, ]
  message(sprintf("[%d/%d] Pulling SEC submissions for %s (%s)", index, nrow(candidate_companies), one_company$symbol, one_company$cik))

  submission_map <- get_sec_submission_map(
    cik = one_company$cik,
    start_year = min(fill_years),
    end_year = max(fill_years),
    headers = SEC_HEADERS
  )

  if (nrow(submission_map) == 0) {
    return(tibble())
  }

  target_rows <- submission_map %>%
    filter(fiscal_year %in% fill_years)

  purrr::map_dfr(seq_len(nrow(target_rows)), function(row_index) {
    Sys.sleep(0.25)
    extract_one_edgar_row(
      cik = one_company$cik,
      symbol = one_company$symbol,
      fiscal_year = target_rows$fiscal_year[row_index],
      filing_info = target_rows[row_index, ],
      headers = SEC_HEADERS
    )
  })
})

if (nrow(live_rows) > 0) {
  edgar_cached <- bind_rows(
    edgar_cached %>%
      transmute(
        cik,
        cached_symbol = symbol,
        year,
        e_score,
        s_score,
        g_score,
        esg_total,
        e_count,
        s_count,
        g_count,
        word_count,
        filing_date
      ),
    live_rows %>%
      transmute(
        cik,
        cached_symbol = symbol,
        year,
        e_score,
        s_score,
        g_score,
        esg_total,
        e_count,
        s_count,
        g_count,
        word_count,
        filing_date
      )
  ) %>%
    arrange(cik, year) %>%
    distinct(cik, year, .keep_all = TRUE)
```

## Assemble the Final Price + ESG Panel

Construct the firm-year panel by crossing each company with every year in the
study window, then left-joining annual price features and EDGAR ESG signals.
Row status flags (`panel_row_status`) track which data sources are available
for each firm-year, making coverage measurement straightforward.

```{r final-panel}
# ---- Start with company × year Cartesian product as the panel base ----
panel_base <- company_master %>%
  select(
    symbol,
    symbol_raw,
    company_name,
    cik,
    gics_sector,
    gics_sub_industry,
    headquarters_location,
    date_added,
    founded,
    snapshot_name,
    snapshot_sector,
    snapshot_industry,
    snapshot_total_esg_risk,
    snapshot_environment_risk,
    snapshot_governance_risk,
    snapshot_social_risk,
    snapshot_controversy_level,
    snapshot_controversy_score,
    snapshot_esg_percentile,
    snapshot_esg_level,
    snapshot_available
  ) %>%
  tidyr::crossing(year = seq.int(START_YEAR, END_YEAR))

# ---- Align cached EDGAR data to the panel base via CIK + year ----
edgar_yearly <- panel_base %>%
  distinct(symbol, cik, year) %>%
  left_join(edgar_cached, by = c("cik", "year")) %>%
  mutate(
    edgar_symbol = coalesce(cached_symbol, symbol),
    edgar_available = !is.na(esg_total)
  ) %>%
  select(
    symbol,
    cik,
    year,
    edgar_symbol,
    edgar_available,
    edgar_e_score = e_score,
    edgar_s_score = s_score,
    edgar_g_score = g_score,
    edgar_esg_total = esg_total,
    edgar_e_count = e_count,
    edgar_s_count = s_count,
    edgar_g_count = g_count,
    edgar_word_count = word_count,
    edgar_filing_date = filing_date
  )

data.table::fwrite(edgar_yearly, OUTPUT_FILES$edgar_yearly)

# ---- Left-join price features and EDGAR signals to complete the panel ----
final_panel <- panel_base %>%
  left_join(
    annual_price_features,
    by = c("symbol", "company_name", "cik", "gics_sector", "gics_sub_industry", "year")
  ) %>%
  left_join(edgar_yearly, by = c("symbol", "cik", "year")) %>%
  mutate(
    price_available = !is.na(end_price),
    edgar_available = coalesce(edgar_available, FALSE),
    # Firms are "post index addition" from the year they joined the S&P 500 onward.
    # Firms missing date_added are treated as always in the index (conservative).
    is_post_index_addition = if_else(!is.na(date_added), year >= lubridate::year(date_added), TRUE),
    panel_row_status = case_when(
      price_available & edgar_available ~ "price_and_edgar",
      price_available & !edgar_available ~ "price_only",
      !price_available & edgar_available ~ "edgar_only",
      TRUE ~ "missing_both"
    )
  ) %>%
  arrange(symbol, year)

data.table::fwrite(final_panel, OUTPUT_FILES$panel)
```

## Coverage Checks

Basic coverage diagnostics: how many firms have price, EDGAR ESG, or snapshot ESG
data in each year? These summaries feed the downstream usability audit and help
spot data gaps before modeling.

```{r coverage}
# ---- Yearly coverage: fraction of firms with each data type ----
coverage_by_year <- final_panel %>%
  group_by(year) %>%
  summarise(
    firms_total = n(),
    firms_with_price = sum(price_available, na.rm = TRUE),
    firms_with_edgar = sum(edgar_available, na.rm = TRUE),
    firms_with_snapshot = sum(snapshot_available, na.rm = TRUE),
    firms_with_price_and_edgar = sum(price_available & edgar_available, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    price_coverage = round(firms_with_price / firms_total, 4),
    edgar_coverage = round(firms_with_edgar / firms_total, 4),
    snapshot_coverage = round(firms_with_snapshot / firms_total, 4)
  )

coverage_overview <- tibble(
  metric = c(
    "constituent_rows",
    "unique_cik",
    "years_in_panel",
    "panel_rows",
    "firms_with_any_price",
    "firms_with_any_edgar",
    "firms_with_snapshot_esg",
    "panel_start_year",
    "panel_end_year"
  ),
  value = c(
    nrow(company_master),
    n_distinct(company_master$cik),
    n_distinct(final_panel$year),
    nrow(final_panel),
    n_distinct(final_panel$symbol[final_panel$price_available]),
    n_distinct(final_panel$cik[final_panel$edgar_available]),
    sum(company_master$snapshot_available, na.rm = TRUE),
    START_YEAR,
    END_YEAR
  )
)

# ---- Company-level availability flags: which firms have data in any year? ----
company_coverage_flags <- final_panel %>%
  group_by(symbol, company_name, cik, gics_sector, gics_sub_industry) %>%
  summarise(
    any_price = any(price_available, na.rm = TRUE),
    any_edgar = any(edgar_available, na.rm = TRUE),
    snapshot_available = any(snapshot_available, na.rm = TRUE),
    first_price_year = if (any(price_available, na.rm = TRUE)) min(year[price_available], na.rm = TRUE) else NA_integer_,
    last_price_year = if (any(price_available, na.rm = TRUE)) max(year[price_available], na.rm = TRUE) else NA_integer_,
    first_edgar_year = if (any(edgar_available, na.rm = TRUE)) min(year[edgar_available], na.rm = TRUE) else NA_integer_,
    last_edgar_year = if (any(edgar_available, na.rm = TRUE)) max(year[edgar_available], na.rm = TRUE) else NA_integer_,
    .groups = "drop"
  )

data.table::fwrite(coverage_by_year, OUTPUT_FILES$coverage_year)
data.table::fwrite(coverage_overview, OUTPUT_FILES$coverage_overview)
data.table::fwrite(company_coverage_flags, OUTPUT_FILES$coverage_company)

coverage_overview %>% knitr::kable()
coverage_by_year %>% knitr::kable()
```


## Output Files

```{r outputs, results='asis'}
output_table <- tibble(
  file = to_project_relative(unname(c(unlist(OUTPUT_FILES), unlist(OUTPUT_USABILITY_FILES)))),
  description = c(
    "Current S&P 500 constituent master table with CIK, sector, and snapshot ESG",
    "Raw cached table of current S&P 500 constituents from Wikipedia",
    "Filtered daily price history for the S&P 500 universe",
    "Annual price features aggregated from daily prices",
    "Annual EDGAR ESG signal table aligned to the current S&P 500 universe",
    "Final annual price + ESG panel",
    "Coverage statistics by year",
    "Overall coverage overview",
    "Firm-level price/EDGAR coverage flags",
    "Dual-denominator usability overview",
    "Dual-denominator coverage by year",
    "Dual-denominator coverage by sector",
    "Field missingness summary for the main panel",
    "Coverage distribution by firm-level usability bucket",
    "Special sample and boundary-case summary",
    "Line chart of dual-denominator coverage by year",
    "Yearly bar chart for jointly available samples",
    "Sector coverage heatmap",
    "Field missingness bar chart",
    "Firm coverage bucket distribution"
  )
)

knitr::kable(output_table)
```

## Notes

This section completes the reproducible data-foundation workflow:

- Pulls the current S&P 500 constituent universe from Wikipedia.
- Reuses the existing Yahoo Finance history cache to build daily prices and annual features.
- Preserves the full cross-sectional ESG snapshot from `SP 500 ESG Risk Ratings.csv`.
- Adds time-varying ESG signals from the pre-built EDGAR 10-K text cache.

The panel `data/derived/sp500_price_esg_panel_2012_2023.csv` is the starting point for
downstream work. The remaining sections of this notebook extend it with:

1. **Data usability audits**: dual-denominator coverage analysis and missingness diagnostics.
2. **Crash-risk labels**: `NCSKEW`, `CRASH_DUMMY`, and related metrics computed from weekly return residuals.
3. **One-year and two-year machine learning warning systems**: full train/validate/test pipelines.

## Data Usability Analysis

The usability audit measures how complete the panel is under two denominator definitions:
- **Full sample** (503 firms): all current constituents, regardless of when they joined the S&P 500
- **Effective post-entry sample**: only firm-years after the company joined the index

This dual view matters because a firm that joined the S&P 500 in 2020 cannot
reasonably be expected to have historical data in 2012.

### Prepare Usability Audit Data

```{r usability-setup}
# ---- Reload the merged panel with strict type coercion ----
panel_for_usability <- readr::read_csv(
  OUTPUT_FILES$panel,
  show_col_types = FALSE
) %>%
  mutate(
    year = as.integer(year),
    date_added = suppressWarnings(as.Date(date_added)),
    price_available = coalesce(as.logical(price_available), FALSE),
    snapshot_available = coalesce(as.logical(snapshot_available), FALSE),
    edgar_available = coalesce(as.logical(edgar_available), FALSE),
    eligible_by_year = coalesce(as.logical(is_post_index_addition), TRUE),
    price_and_snapshot = price_available & snapshot_available,
    price_and_edgar = price_available & edgar_available,
    has_cik = !is.na(cik) & stringr::str_trim(as.character(cik)) != "",
    has_date_added = !is.na(date_added),
    sector_standardized = dplyr::case_when(
      gics_sector == "Technology" ~ "Information Technology",
      gics_sector == "Healthcare" ~ "Health Care",
      gics_sector == "Financial Services" ~ "Financials",
      TRUE ~ gics_sector
    )
  )

legacy_coverage_by_year <- readr::read_csv(
  OUTPUT_FILES$coverage_year,
  show_col_types = FALSE
)

company_flags_for_usability <- readr::read_csv(
  OUTPUT_FILES$coverage_company,
  show_col_types = FALSE
) %>%
  mutate(
    any_price = coalesce(as.logical(any_price), FALSE),
    any_edgar = coalesce(as.logical(any_edgar), FALSE),
    snapshot_available = coalesce(as.logical(snapshot_available), FALSE),
    has_cik = !is.na(cik) & stringr::str_trim(as.character(cik)) != ""
  )

# ---- Define the five coverage metrics tracked throughout the audit ----
# "available" = the data source exists for that firm-year row
# Joint metrics (price_and_snapshot, price_and_edgar) measure co-occurrence
metric_levels <- c(
  "price_available",
  "snapshot_available",
  "edgar_available",
  "price_and_snapshot",
  "price_and_edgar"
)

metric_labels <- c(
  price_available = "Price Data",
  snapshot_available = "Snapshot ESG",
  edgar_available = "EDGAR ESG",
  price_and_snapshot = "Price + Snapshot ESG",
  price_and_edgar = "Price + EDGAR ESG"
)

denominator_labels <- c(
  full_sample = "Full Sample (503 Firms)",
  effective_sample = "Effective Post-Entry Sample"
)

metric_colors <- c(
  "Price Data" = "#2F6DB3",
  "Snapshot ESG" = "#2A9D55",
  "EDGAR ESG" = "#E68613",
  "Price + Snapshot ESG" = "#1F8A8A",
  "Price + EDGAR ESG" = "#4D4D4D"
)

bucket_levels <- c(
  "price+snapshot+edgar",
  "price+snapshot",
  "price+edgar",
  "price_only",
  "snapshot_only",
  "none"
)

# ---- Classify each panel column into a logical field group for the missingness audit ----
classify_panel_field_group <- function(field_name) {
  dplyr::case_when(
    field_name %in% c("symbol", "symbol_raw", "company_name", "cik", "gics_sector",
                      "gics_sub_industry", "headquarters_location", "date_added",
                      "founded", "year") ~ "company_static",
    stringr::str_starts(field_name, "snapshot_") ~ "snapshot_esg",
    field_name %in% c("n_trading_days", "first_trade", "last_trade", "start_price",
                      "end_price", "annual_return", "annual_log_return",
                      "annual_volatility", "avg_volume", "median_volume",
                      "max_drawdown") ~ "price_features",
    stringr::str_starts(field_name, "edgar_") ~ "edgar_esg",
    field_name %in% c("price_available", "eligible_by_year", "price_and_snapshot",
                      "price_and_edgar", "has_cik", "has_date_added",
                      "is_post_index_addition", "panel_row_status",
                      "sector_standardized") ~ "flags",
    TRUE ~ "other"
  )
}

# The full-sample denominator is always 503 (or the total distinct symbols in the panel)
full_sample_denominator <- dplyr::n_distinct(panel_for_usability$symbol)
```

### Coverage Summary and CSV Export

The dual-denominator coverage tables are built by pivoting the five availability
indicators into long form and computing rates under both denominator definitions.

```{r usability-tables}
# ---- Pivot availability indicators to long format for grouped aggregation ----
firm_year_metric_long <- panel_for_usability %>%
  select(
    year,
    symbol,
    sector_standardized,
    eligible_by_year,
    all_of(metric_levels)
  ) %>%
  pivot_longer(
    cols = all_of(metric_levels),
    names_to = "metric",
    values_to = "available"
  ) %>%
  mutate(
    available = coalesce(as.logical(available), FALSE),
    metric = factor(metric, levels = metric_levels, ordered = TRUE),
    metric_label = unname(metric_labels[as.character(metric)])
  )

# ---- Full-sample coverage: denominator = all 503 firms, every year ----
coverage_by_year_full <- firm_year_metric_long %>%
  group_by(year, metric, metric_label) %>%
  summarise(
    numerator = sum(available, na.rm = TRUE),
    denominator = full_sample_denominator,
    coverage_rate = numerator / denominator,
    denominator_type = "full_sample",
    .groups = "drop"
  )

# ---- Effective-sample coverage: denominator = only post-index-entry firm-years ----
coverage_by_year_effective <- firm_year_metric_long %>%
  filter(eligible_by_year) %>%
  group_by(year, metric, metric_label) %>%
  summarise(
    numerator = sum(available, na.rm = TRUE),
    denominator = n(),
    coverage_rate = if_else(denominator > 0, numerator / denominator, NA_real_),
    denominator_type = "effective_sample",
    .groups = "drop"
  )

# ---- Combine full-sample and effective-sample into one dual-denominator table ----
coverage_by_year_dual <- bind_rows(
  coverage_by_year_full,
  coverage_by_year_effective
) %>%
  mutate(
    denominator_label = unname(denominator_labels[denominator_type])
  ) %>%
  arrange(denominator_type, year, metric)

overall_coverage_dual <- bind_rows(
  firm_year_metric_long %>%
    group_by(metric, metric_label) %>%
    summarise(
      numerator = sum(available, na.rm = TRUE),
      denominator = n(),
      coverage_rate = numerator / denominator,
      denominator_type = "full_sample",
      .groups = "drop"
    ),
  firm_year_metric_long %>%
    filter(eligible_by_year) %>%
    group_by(metric, metric_label) %>%
    summarise(
      numerator = sum(available, na.rm = TRUE),
      denominator = n(),
      coverage_rate = if_else(denominator > 0, numerator / denominator, NA_real_),
      denominator_type = "effective_sample",
      .groups = "drop"
    )
) %>%
  mutate(
    summary_group = "coverage",
    denominator_label = unname(denominator_labels[denominator_type])
  )

overview_base <- tibble(
  summary_group = "base",
  denominator_type = "not_applicable",
  denominator_label = "Not Applicable",
  metric = c(
    "constituent_rows",
    "years_in_panel",
    "panel_rows",
    "firms_with_any_price",
    "firms_with_any_snapshot",
    "firms_with_any_edgar",
    "companies_missing_date_added",
    "companies_missing_cik"
  ),
  metric_label = c(
    "Number of Firms",
    "Number of Years",
    "Firm-Year Rows",
    "Firms with Price Data in At Least One Year",
    "Firms with Snapshot ESG",
    "Firms with EDGAR ESG in At Least One Year",
    "Firms Missing date_added",
    "Firms Missing CIK"
  ),
  numerator = c(
    dplyr::n_distinct(panel_for_usability$symbol),
    dplyr::n_distinct(panel_for_usability$year),
    nrow(panel_for_usability),
    dplyr::n_distinct(panel_for_usability$symbol[panel_for_usability$price_available]),
    dplyr::n_distinct(panel_for_usability$symbol[panel_for_usability$snapshot_available]),
    dplyr::n_distinct(panel_for_usability$symbol[panel_for_usability$edgar_available]),
    dplyr::n_distinct(panel_for_usability$symbol[!panel_for_usability$has_date_added]),
    dplyr::n_distinct(panel_for_usability$symbol[!panel_for_usability$has_cik])
  ),
  denominator = NA_real_,
  coverage_rate = NA_real_
)

overview_dual_denominator <- bind_rows(
  overview_base,
  overall_coverage_dual
) %>%
  arrange(summary_group, denominator_type, metric)

coverage_by_sector_dual <- bind_rows(
  firm_year_metric_long %>%
    group_by(sector_standardized, metric, metric_label) %>%
    summarise(
      firms_in_sector = dplyr::n_distinct(symbol),
      numerator = sum(available, na.rm = TRUE),
      denominator = n(),
      coverage_rate = if_else(denominator > 0, numerator / denominator, NA_real_),
      denominator_type = "full_sample",
      coverage_period = "all_years_avg",
      .groups = "drop"
    ),
  firm_year_metric_long %>%
    filter(eligible_by_year) %>%
    group_by(sector_standardized, metric, metric_label) %>%
    summarise(
      firms_in_sector = dplyr::n_distinct(symbol),
      numerator = sum(available, na.rm = TRUE),
      denominator = n(),
      coverage_rate = if_else(denominator > 0, numerator / denominator, NA_real_),
      denominator_type = "effective_sample",
      coverage_period = "all_years_avg",
      .groups = "drop"
    ),
  firm_year_metric_long %>%
    filter(year == END_YEAR) %>%
    group_by(sector_standardized, metric, metric_label) %>%
    summarise(
      firms_in_sector = dplyr::n_distinct(symbol),
      numerator = sum(available, na.rm = TRUE),
      denominator = n(),
      coverage_rate = if_else(denominator > 0, numerator / denominator, NA_real_),
      denominator_type = "full_sample",
      coverage_period = sprintf("year_%s", END_YEAR),
      .groups = "drop"
    ),
  firm_year_metric_long %>%
    filter(year == END_YEAR, eligible_by_year) %>%
    group_by(sector_standardized, metric, metric_label) %>%
    summarise(
      firms_in_sector = dplyr::n_distinct(symbol),
      numerator = sum(available, na.rm = TRUE),
      denominator = n(),
      coverage_rate = if_else(denominator > 0, numerator / denominator, NA_real_),
      denominator_type = "effective_sample",
      coverage_period = sprintf("year_%s", END_YEAR),
      .groups = "drop"
    )
) %>%
  mutate(
    denominator_label = unname(denominator_labels[denominator_type])
  ) %>%
  arrange(coverage_period, denominator_type, sector_standardized, metric)

# ---- Field-level missingness: how many NAs per column? ----
field_missingness <- tibble(
  field = names(panel_for_usability),
  missing_count = purrr::map_int(names(panel_for_usability), ~ sum(is.na(panel_for_usability[[.x]]))),
  missing_rate = purrr::map_dbl(names(panel_for_usability), ~ mean(is.na(panel_for_usability[[.x]])))
) %>%
  mutate(
    non_missing_rate = 1 - missing_rate,
    field_group = purrr::map_chr(field, classify_panel_field_group)
  ) %>%
  arrange(desc(missing_rate), field)

# ---- Classify each company into one of six fixed coverage buckets ----
company_coverage_buckets <- company_flags_for_usability %>%
  mutate(
    company_bucket = case_when(
      any_price & snapshot_available & any_edgar ~ "price+snapshot+edgar",
      any_price & snapshot_available & !any_edgar ~ "price+snapshot",
      any_price & !snapshot_available & any_edgar ~ "price+edgar",
      any_price & !snapshot_available & !any_edgar ~ "price_only",
      !any_price & snapshot_available & !any_edgar ~ "snapshot_only",
      TRUE ~ "none"
    ),
    company_bucket = factor(company_bucket, levels = bucket_levels, ordered = TRUE)
  ) %>%
  count(company_bucket, name = "company_count", .drop = FALSE) %>%
  mutate(
    share = company_count / sum(company_count),
    company_bucket = as.character(company_bucket)
  )

# ---- Document boundary cases: missing identifiers, edge firms, ineligible rows ----
special_cases_summary <- tibble(
  scope = c(
    "company",
    "company",
    "company",
    "company",
    "company",
    "company",
    "company",
    "firm_year"
  ),
  metric = c(
    "companies_missing_date_added",
    "companies_missing_cik",
    "companies_without_any_price",
    "companies_without_any_snapshot",
    "companies_without_any_edgar",
    "companies_with_edgar_but_no_price",
    "companies_bucketed_as_none",
    "firm_years_not_eligible_by_year"
  ),
  value = c(
    dplyr::n_distinct(panel_for_usability$symbol[!panel_for_usability$has_date_added]),
    dplyr::n_distinct(panel_for_usability$symbol[!panel_for_usability$has_cik]),
    sum(!company_flags_for_usability$any_price, na.rm = TRUE),
    sum(!company_flags_for_usability$snapshot_available, na.rm = TRUE),
    sum(!company_flags_for_usability$any_edgar, na.rm = TRUE),
    sum(company_flags_for_usability$any_edgar & !company_flags_for_usability$any_price, na.rm = TRUE),
    company_coverage_buckets$company_count[company_coverage_buckets$company_bucket == "none"],
    sum(!panel_for_usability$eligible_by_year, na.rm = TRUE)
  ),
  note = c(
    "Firms missing date_added are retained in the effective-sample denominator by default",
    "Missing CIK affects the interpretation of downstream EDGAR matching",
    "No company-level observation ever has price_available = TRUE",
    "No company-level observation ever has snapshot_available = TRUE",
    "No company-level observation ever has edgar_available = TRUE",
    "These firms are folded into the none bucket",
    "Fallback category within the fixed six coverage buckets",
    "These observations enter the denominator only under the full_sample definition"
  )
)

data.table::fwrite(overview_dual_denominator, OUTPUT_USABILITY_FILES$overview)
data.table::fwrite(coverage_by_year_dual, OUTPUT_USABILITY_FILES$coverage_year)
data.table::fwrite(coverage_by_sector_dual, OUTPUT_USABILITY_FILES$coverage_sector)
data.table::fwrite(field_missingness, OUTPUT_USABILITY_FILES$field_missingness)
data.table::fwrite(company_coverage_buckets, OUTPUT_USABILITY_FILES$company_buckets)
data.table::fwrite(special_cases_summary, OUTPUT_USABILITY_FILES$special_cases)

overview_dual_denominator %>%
  filter(summary_group == "base") %>%
  select(metric_label, numerator) %>%
  knitr::kable()

coverage_by_year_dual %>%
  filter(metric %in% c("price_available", "snapshot_available", "edgar_available")) %>%
  select(year, denominator_label, metric_label, numerator, denominator, coverage_rate) %>%
  head(12) %>%
  knitr::kable()
```

### Visualization Outputs

```{r usability-plots, fig.width=10, fig.height=6}
yearly_coverage_plot <- coverage_by_year_dual %>%
  ggplot(aes(x = year, y = coverage_rate, color = metric_label, group = metric_label)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.8) +
  facet_wrap(~ denominator_label, ncol = 1) +
  scale_color_manual(values = metric_colors) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "S&P 500 Coverage Rates (Dual Denominator)",
    subtitle = "Showing both the full sample of 503 firms and the effective post-entry sample",
    x = "Year",
    y = "Coverage Rate",
    color = "Data Type"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")

joint_coverage_bar_plot <- coverage_by_year_dual %>%
  filter(
    denominator_type == "effective_sample",
    metric %in% c("price_and_snapshot", "price_and_edgar")
  ) %>%
  ggplot(aes(x = factor(year), y = coverage_rate, fill = metric_label)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_manual(values = metric_colors[c("Price + Snapshot ESG", "Price + EDGAR ESG")]) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Coverage of Jointly Available Samples",
    subtitle = "Effective post-entry sample definition",
    x = "Year",
    y = "Coverage Rate",
    fill = "Joint Sample Type"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")

sector_heatmap_data <- coverage_by_sector_dual %>%
  filter(
    denominator_type == "effective_sample",
    coverage_period == "all_years_avg"
  ) %>%
  group_by(sector_standardized) %>%
  mutate(
    edgar_sort = coverage_rate[metric == "edgar_available"][1]
  ) %>%
  ungroup() %>%
  mutate(
    sector_standardized = reorder(sector_standardized, edgar_sort, na.rm = TRUE),
    metric_label = factor(metric_label, levels = unname(metric_labels), ordered = TRUE)
  )

sector_heatmap_plot <- sector_heatmap_data %>%
  ggplot(aes(x = metric_label, y = sector_standardized, fill = coverage_rate)) +
  geom_tile(color = "white", linewidth = 0.3) +
  geom_text(aes(label = ifelse(is.na(coverage_rate), "NA", scales::percent(coverage_rate, accuracy = 1))), size = 3) +
  scale_fill_gradient(low = "#F1F5F9", high = "#2F6DB3", labels = scales::percent_format(accuracy = 1), na.value = "#D9D9D9") +
  labs(
    title = "Sector-Level Coverage Heatmap",
    subtitle = "Effective post-entry sample definition using average coverage across the full period",
    x = "Data Type",
    y = "Standardized Sector",
    fill = "Coverage Rate"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

field_missingness_plot <- field_missingness %>%
  mutate(field = reorder(field, missing_rate)) %>%
  ggplot(aes(x = field, y = missing_rate, fill = field_group)) +
  geom_col(width = 0.75) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Main Panel Field Missingness",
    subtitle = "Missingness rates grouped by field family",
    x = "Field",
    y = "Missing Rate",
    fill = "Field Group"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.y = element_text(size = 9)) +
  coord_flip()

bucket_plot <- company_coverage_buckets %>%
  mutate(company_bucket = factor(company_bucket, levels = bucket_levels, ordered = TRUE)) %>%
  ggplot(aes(x = company_bucket, y = company_count, fill = company_bucket)) +
  geom_col(width = 0.7, show.legend = FALSE) +
  geom_text(aes(label = scales::percent(share, accuracy = 0.1)), vjust = -0.4, size = 3.5) +
  labs(
    title = "Firm Coverage Bucket Distribution",
    subtitle = "Main usability combinations aggregated at the firm level",
    x = "Coverage Bucket",
    y = "Number of Firms"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 18, hjust = 1))

ggplot2::ggsave(OUTPUT_USABILITY_FILES$yearly_lines, yearly_coverage_plot, width = 10, height = 7, dpi = 160)
ggplot2::ggsave(OUTPUT_USABILITY_FILES$joint_bars, joint_coverage_bar_plot, width = 10, height = 6, dpi = 160)
ggplot2::ggsave(OUTPUT_USABILITY_FILES$sector_heatmap, sector_heatmap_plot, width = 11, height = 7.5, dpi = 160)
ggplot2::ggsave(OUTPUT_USABILITY_FILES$missingness_bar, field_missingness_plot, width = 11, height = 10, dpi = 160)
ggplot2::ggsave(OUTPUT_USABILITY_FILES$bucket_bar, bucket_plot, width = 10, height = 6, dpi = 160)

yearly_coverage_plot
```

### Audit Validation Checks

Assert invariants that must hold for the audit tables to be internally consistent:
- Denominators match expected counts
- Coverage rates are bounded in [0, 1]
- Joint coverage never exceeds the min of its components
- Snapshot coverage is constant across years (cross-sectional source)
- Sector labels have been standardized
- New dual-denominator definitions match legacy single-denominator figures

```{r usability-validation}
# ---- Cross-check new coverage definitions against legacy outputs ----
legacy_compare <- coverage_by_year_dual %>%
  filter(
    denominator_type == "full_sample",
    metric %in% c("price_available", "snapshot_available", "edgar_available")
  ) %>%
  select(year, metric, coverage_rate) %>%
  pivot_wider(names_from = metric, values_from = coverage_rate) %>%
  left_join(
    legacy_coverage_by_year %>%
      select(year, price_coverage, snapshot_coverage, edgar_coverage),
    by = "year"
  )

joint_edgar_check <- coverage_by_year_dual %>%
  select(year, denominator_type, metric, coverage_rate) %>%
  pivot_wider(names_from = metric, values_from = coverage_rate) %>%
  transmute(check = price_and_edgar <= pmin(price_available, edgar_available)) %>%
  pull(check)

joint_snapshot_check <- coverage_by_year_dual %>%
  select(year, denominator_type, metric, coverage_rate) %>%
  pivot_wider(names_from = metric, values_from = coverage_rate) %>%
  transmute(check = price_and_snapshot <= pmin(price_available, snapshot_available)) %>%
  pull(check)

stopifnot(all(coverage_by_year_dual$denominator[coverage_by_year_dual$denominator_type == "full_sample"] == 503))
stopifnot(all(coverage_by_year_dual$denominator[coverage_by_year_dual$denominator_type == "effective_sample"] <= 503))
stopifnot(all(coverage_by_year_dual$coverage_rate[!is.na(coverage_by_year_dual$coverage_rate)] >= 0))
stopifnot(all(coverage_by_year_dual$coverage_rate[!is.na(coverage_by_year_dual$coverage_rate)] <= 1))
stopifnot(all(coverage_by_year_dual$coverage_rate[!is.na(coverage_by_year_dual$coverage_rate)] %>% is.finite()))
stopifnot(all(joint_edgar_check[!is.na(joint_edgar_check)]))
stopifnot(all(joint_snapshot_check[!is.na(joint_snapshot_check)]))
stopifnot(
  dplyr::n_distinct(
    coverage_by_year_dual$coverage_rate[
      coverage_by_year_dual$denominator_type == "full_sample" &
        coverage_by_year_dual$metric == "snapshot_available"
    ]
  ) == 1
)
stopifnot(!any(panel_for_usability$sector_standardized %in% c("Technology", "Healthcare", "Financial Services"), na.rm = TRUE))
stopifnot(all(round(legacy_compare$price_available, 4) == round(legacy_compare$price_coverage, 4)))
stopifnot(all(round(legacy_compare$snapshot_available, 4) == round(legacy_compare$snapshot_coverage, 4)))
stopifnot(all(round(legacy_compare$edgar_available, 4) == round(legacy_compare$edgar_coverage, 4)))

validation_summary <- tibble(
  check = c(
    "Full-sample denominator is always 503",
    "Effective-sample denominator never exceeds 503",
    "All coverage rates lie within [0, 1]",
    "Joint coverage never exceeds single-source coverage",
    "Snapshot annual coverage is constant",
    "Sector labels are standardized",
    "New and legacy yearly coverage definitions match"
  ),
  result = "PASS"
)

knitr::kable(validation_summary)
```

### Usability Analysis Output Files

```{r usability-output-files}
usability_output_table <- tibble(
  file = to_project_relative(unname(unlist(OUTPUT_USABILITY_FILES))),
  description = c(
    "Dual-denominator usability overview",
    "Dual-denominator coverage by year",
    "Dual-denominator coverage by sector",
    "Field missingness summary for the main panel",
    "Coverage distribution by firm-level usability bucket",
    "Special sample and boundary-case summary",
    "Line chart of dual-denominator coverage by year",
    "Yearly bar chart for jointly available samples",
    "Sector coverage heatmap",
    "Field missingness bar chart",
    "Firm coverage bucket distribution"
  )
)

knitr::kable(usability_output_table)
```

## Machine Learning Helper Functions

Shared functions for crash-label construction, classification metrics, feature
preprocessing, and model fitting. Used by the two-year crash-risk warning pipeline
in the next section.

```{r ml-warning-helpers}
# ---------------------------------------------------------------------------
# Python fallback for Yahoo Finance price downloads
# ---------------------------------------------------------------------------
# When cached price files are missing, we call a Python script that uses
# Yahoo Finance's chart API. R's quantmod/tidyquant can be unreliable for
# bulk downloads, hence the Python fallback.

detect_workspace_python <- function() {
  candidates <- c(
    file.path(Sys.getenv("USERPROFILE"), ".cache", "codex-runtimes", "codex-primary-runtime", "dependencies", "python", "python.exe"),
    Sys.which("python")
  )

  valid_candidates <- candidates[nzchar(candidates) & file.exists(candidates)]
  if (length(valid_candidates) == 0) {
    return(NA_character_)
  }

  valid_candidates[[1]]
}

fetch_yahoo_prices_python <- function(tickers, start_date, end_date_exclusive, output_file) {
  python_path <- detect_workspace_python()

  if (is.na(python_path) || !file.exists(python_path)) {
    stop("No accessible Python runtime was found for the Yahoo Finance fallback.")
  }

  tickers <- unique(tickers[!is.na(tickers) & nzchar(tickers)])
  tickers_file <- tempfile(fileext = ".txt")
  script_file <- tempfile(fileext = ".py")
  writeLines(tickers, tickers_file, useBytes = TRUE)

  script_lines <- c(
    "import csv",
    "import json",
    "import sys",
    "import time",
    "import urllib.parse",
    "import urllib.request",
    "from concurrent.futures import ThreadPoolExecutor, as_completed",
    "from datetime import datetime, timezone",
    "",
    "tickers_file, output_file, start_date, end_date = sys.argv[1:5]",
    "with open(tickers_file, 'r', encoding='utf-8') as handle:",
    "    tickers = [line.strip() for line in handle if line.strip()]",
    "",
    "period1 = int(datetime.fromisoformat(start_date).replace(tzinfo=timezone.utc).timestamp())",
    "period2 = int(datetime.fromisoformat(end_date).replace(tzinfo=timezone.utc).timestamp())",
    "headers = {'User-Agent': 'Mozilla/5.0'}",
    "",
    "def fetch_one(ticker):",
    "    encoded_ticker = urllib.parse.quote(ticker)",
    "    url = f'https://query2.finance.yahoo.com/v8/finance/chart/{encoded_ticker}?period1={period1}&period2={period2}&interval=1d&events=div%2Csplits&includeAdjustedClose=true'",
    "    last_error = None",
    "    for attempt in range(3):",
    "        try:",
    "            request = urllib.request.Request(url, headers=headers)",
    "            with urllib.request.urlopen(request, timeout=30) as response:",
    "                payload = json.loads(response.read().decode('utf-8'))",
    "            result = (payload.get('chart', {}) or {}).get('result')",
    "            if not result:",
    "                return ticker, [], str((payload.get('chart', {}) or {}).get('error'))",
    "            result = result[0]",
    "            timestamps = result.get('timestamp') or []",
    "            indicators = result.get('indicators') or {}",
    "            quote = (indicators.get('quote') or [{}])[0]",
    "            adj_block = (indicators.get('adjclose') or [{}])[0]",
    "            adj_values = adj_block.get('adjclose')",
    "            if adj_values is None:",
    "                adj_values = quote.get('close') or []",
    "            volume_values = quote.get('volume') or []",
    "            rows = []",
    "            for idx, one_timestamp in enumerate(timestamps):",
    "                adj_value = adj_values[idx] if idx < len(adj_values) else None",
    "                volume_value = volume_values[idx] if idx < len(volume_values) else None",
    "                if adj_value is None:",
    "                    continue",
    "                date_value = datetime.utcfromtimestamp(one_timestamp).strftime('%Y-%m-%d')",
    "                rows.append([ticker, date_value, adj_value, '' if volume_value is None else volume_value])",
    "            return ticker, rows, None",
    "        except Exception as error:",
    "            last_error = str(error)",
    "            time.sleep(0.8 * (attempt + 1))",
    "    return ticker, [], last_error",
    "",
    "all_rows = []",
    "failures = []",
    "with ThreadPoolExecutor(max_workers=8) as executor:",
    "    futures = {executor.submit(fetch_one, ticker): ticker for ticker in tickers}",
    "    for future in as_completed(futures):",
    "        ticker, rows, error_message = future.result()",
    "        if rows:",
    "            all_rows.extend(rows)",
    "        else:",
    "            failures.append((ticker, error_message))",
    "",
    "with open(output_file, 'w', newline='', encoding='utf-8') as handle:",
    "    writer = csv.writer(handle)",
    "    writer.writerow(['price_symbol', 'date', 'adj_close', 'volume'])",
    "    writer.writerows(all_rows)",
    "",
    "if failures:",
    "    sys.stderr.write('Failed tickers: ' + str(len(failures)) + '\\n')",
    "    preview = '; '.join([f'{ticker}={message}' for ticker, message in failures[:10]])",
    "    sys.stderr.write(preview + '\\n')",
    "",
    "sys.stdout.write(str(len(all_rows)))"
  )

  writeLines(script_lines, script_file, useBytes = TRUE)

  command_output <- system2(
    command = python_path,
    args = c(script_file, tickers_file, output_file, start_date, end_date_exclusive),
    stdout = TRUE,
    stderr = TRUE
  )

  if (!file.exists(output_file)) {
    stop("The Python Yahoo download fallback did not create the expected price file.")
  }

  command_output
}

# ---------------------------------------------------------------------------
# Crash-risk label construction (following Chen, Hong & Stein 2001 / JFE)
# ---------------------------------------------------------------------------
# Crash metrics are computed from weekly residual returns after removing
# market-wide movements. The key steps:
#   1. Aggregate daily log returns to weekly
#   2. Regress firm weekly returns on lead/lag market returns (Dimson 1979)
#   3. Compute NCSKEW (negative conditional skewness) and CRASH_DUMMY from residuals
# A "crash" week is any residual return below mean - 3.09 * sd (~0.1% tail).

compute_symbol_year_crash_metrics <- function(firm_data) {
  n_obs <- nrow(firm_data)
  if (n_obs < 26) {
    return(NULL)
  }

  firm_sorted <- firm_data %>%
    arrange(week_id) %>%
    mutate(
      mkt_lag2 = lag(ret_market, 2),
      mkt_lag1 = lag(ret_market, 1),
      mkt_lead1 = lead(ret_market, 1),
      mkt_lead2 = lead(ret_market, 2)
    )

  reg_data <- firm_sorted %>%
    filter(
      !is.na(mkt_lag2), !is.na(mkt_lag1),
      !is.na(ret_market), !is.na(mkt_lead1), !is.na(mkt_lead2)
    )

  if (nrow(reg_data) < 20) {
    return(NULL)
  }

  # Regress firm weekly returns on lead/lag market returns to isolate firm-specific residual moves.
  fit <- tryCatch(
    lm(ret_week ~ mkt_lag2 + mkt_lag1 + ret_market + mkt_lead1 + mkt_lead2, data = reg_data),
    error = function(error) NULL
  )

  if (is.null(fit)) {
    return(NULL)
  }

  residual_returns <- residuals(fit)
  transformed_returns <- log(1 + residual_returns)
  transformed_returns <- transformed_returns[is.finite(transformed_returns)]

  n <- length(transformed_returns)
  if (n < 20) {
    return(NULL)
  }

  sum_w3 <- sum(transformed_returns^3)
  sum_w2 <- sum(transformed_returns^2)

  ncskew <- NA_real_
  if (sum_w2 > 0) {
    ncskew <- -(n * (n - 1)^(1.5) * sum_w3) / ((n - 1) * (n - 2) * sum_w2^(1.5))
  }

  w_mean <- mean(transformed_returns, na.rm = TRUE)
  w_sd <- sd(transformed_returns, na.rm = TRUE)
  # A crash year is flagged when at least one residual return breaches the 3.09-standard-deviation cutoff.
  crash_dummy <- as.integer(any(transformed_returns < w_mean - 3.09 * w_sd, na.rm = TRUE))

  tibble(
    NCSKEW = ncskew,
    CRASH_DUMMY = crash_dummy,
    SIGMA = w_sd,
    RET_MEAN = w_mean
  )
}

build_sp500_crash_metrics <- function(price_data) {
  # Build weekly firm and market returns because the crash-risk labels are defined on weekly residuals.
  daily_returns <- price_data %>%
    arrange(symbol, date) %>%
    group_by(symbol) %>%
    mutate(ret_daily = log(adj_close / lag(adj_close))) %>%
    ungroup() %>%
    filter(!is.na(ret_daily), is.finite(ret_daily))

  market_daily <- daily_returns %>%
    group_by(date) %>%
    summarise(
      ret_market = mean(ret_daily, na.rm = TRUE),
      .groups = "drop"
    )

  daily_with_week <- daily_returns %>%
    left_join(market_daily, by = "date") %>%
    mutate(
      year = lubridate::year(date),
      week_id = paste0(year, "_W", sprintf("%02d", lubridate::isoweek(date)))
    )

  weekly_returns <- daily_with_week %>%
    group_by(symbol, year, week_id) %>%
    summarise(
      ret_week = sum(ret_daily, na.rm = TRUE),
      ret_market = sum(ret_market, na.rm = TRUE),
      avg_volume = mean(volume, na.rm = TRUE),
      n_days = n(),
      .groups = "drop"
    ) %>%
    filter(n_days >= 3)

  symbol_year_groups <- weekly_returns %>%
    group_by(symbol, year) %>%
    group_split()

  crash_rows <- lapply(symbol_year_groups, function(group_data) {
    metric_row <- compute_symbol_year_crash_metrics(group_data)
    if (is.null(metric_row)) {
      return(NULL)
    }

    bind_cols(
      tibble(symbol = group_data$symbol[[1]], year = group_data$year[[1]]),
      metric_row
    )
  })

  crash_metrics_df <- bind_rows(Filter(Negate(is.null), crash_rows))

  monthly_volume <- daily_returns %>%
    mutate(
      year = lubridate::year(date),
      month = lubridate::month(date)
    ) %>%
    group_by(symbol, year, month) %>%
    summarise(
      monthly_volume = sum(volume, na.rm = TRUE),
      .groups = "drop"
    )

  annual_volume <- monthly_volume %>%
    group_by(symbol, year) %>%
    summarise(
      avg_monthly_volume = mean(monthly_volume, na.rm = TRUE),
      .groups = "drop"
    )

  # DTURN captures the year-over-year change in average monthly trading volume.
  dturn_df <- annual_volume %>%
    arrange(symbol, year) %>%
    group_by(symbol) %>%
    mutate(DTURN = avg_monthly_volume - lag(avg_monthly_volume)) %>%
    ungroup() %>%
    select(symbol, year, DTURN)

  crash_metrics_df %>%
    left_join(dturn_df, by = c("symbol", "year")) %>%
    arrange(symbol, year)
}

# ---------------------------------------------------------------------------
# Classification metrics (all handle NA values via valid_index filtering)
# ---------------------------------------------------------------------------
# ROC AUC computed via the Wilcoxon-Mann-Whitney rank-sum formula

calc_roc_auc <- function(truth, probability) {
  truth <- as.integer(truth)
  valid_index <- !is.na(truth) & !is.na(probability)
  truth <- truth[valid_index]
  probability <- probability[valid_index]

  positive_total <- sum(truth == 1)
  negative_total <- sum(truth == 0)

  if (positive_total == 0 || negative_total == 0) {
    return(NA_real_)
  }

  ranks <- rank(probability, ties.method = "average")
  positive_rank_sum <- sum(ranks[truth == 1])

  (positive_rank_sum - positive_total * (positive_total + 1) / 2) / (positive_total * negative_total)
}

# PR AUC via the trapezoidal rule on sorted predictions
calc_pr_auc <- function(truth, probability) {
  truth <- as.integer(truth)
  valid_index <- !is.na(truth) & !is.na(probability)
  truth <- truth[valid_index]
  probability <- probability[valid_index]

  positive_total <- sum(truth == 1)
  if (positive_total == 0) {
    return(NA_real_)
  }

  order_index <- order(probability, decreasing = TRUE)
  sorted_truth <- truth[order_index]
   true_positive <- cumsum(sorted_truth == 1)
  false_positive <- cumsum(sorted_truth == 0)
  precision <- true_positive / (true_positive + false_positive)
  recall <- true_positive / positive_total
  precision <- c(1, precision)
  recall <- c(0, recall)

  sum((recall[-1] - recall[-length(recall)]) * precision[-1], na.rm = TRUE)
}

# F-beta score with beta = 0.5 (weights precision 2× more than recall).
# We prioritize precision because flagging too many non-crash firms as "high risk"
# erodes the credibility of the warning system more than missing a few crashes.
calc_fbeta <- function(precision, recall, beta_weight = 0.5) {
  if (is.na(precision) || is.na(recall) || (precision + recall) == 0) {
    return(NA_real_)
  }

  beta_square <- beta_weight ^ 2
  (1 + beta_square) * precision * recall / (beta_square * precision + recall)
}

calc_precision_at_top_share <- function(truth, probability, share = 0.10) {
  truth <- as.integer(truth)
  valid_index <- !is.na(truth) & !is.na(probability)
  truth <- truth[valid_index]
  probability <- probability[valid_index]

  if (length(truth) == 0) {
    return(NA_real_)
  }

  k <- max(1, ceiling(length(truth) * share))
  order_index <- order(probability, decreasing = TRUE)
  mean(truth[order_index[seq_len(k)]] == 1)
}

calc_top_n_hit_summary <- function(truth, probability, top_n = 20) {
  truth <- as.integer(truth)
  valid_index <- !is.na(truth) & !is.na(probability)
  truth <- truth[valid_index]
  probability <- probability[valid_index]

  if (length(truth) == 0) {
    return(list(rate = NA_real_, hits = NA_integer_, k = 0L))
  }

  k <- min(top_n, length(truth))
  order_index <- order(probability, decreasing = TRUE)
  hits <- sum(truth[order_index[seq_len(k)]] == 1)

  list(
    rate = hits / k,
    hits = hits,
    k = k
  )
}

build_pr_curve_data <- function(actual, probability, model_label) {
  valid_index <- !is.na(actual) & !is.na(probability)
  actual <- as.integer(actual[valid_index])
  probability <- as.numeric(probability[valid_index])

  if (length(actual) == 0 || sum(actual == 1) == 0) {
    return(tibble())
  }

  order_index <- order(probability, decreasing = TRUE)
  sorted_truth <- actual[order_index]
  sorted_probability <- probability[order_index]
  true_positive <- cumsum(sorted_truth == 1)
  false_positive <- cumsum(sorted_truth == 0)
  positive_total <- sum(sorted_truth == 1)

  tibble(
    model_label = model_label,
    step = seq_along(sorted_probability),
    threshold = sorted_probability,
    precision = true_positive / (true_positive + false_positive),
    recall = true_positive / positive_total
  )
}

choose_threshold_from_validation <- function(threshold_table) {
  threshold_table %>%
    mutate(
      rank_f0_5 = dplyr::dense_rank(desc(mean_f0_5)),
      rank_precision = dplyr::dense_rank(desc(mean_precision))
    ) %>%
    arrange(rank_f0_5, rank_precision, desc(mean_pr_auc_proxy), threshold) %>%
    slice(1) %>%
    select(-rank_f0_5, -rank_precision)
}

# ---- Feature preprocessing ----
# Prepares feature matrices for model fitting:
#   - Continuous features: median-impute NAs, optionally standardize (z-score)
#   - Binary features (missingness flags, availability dummies): impute NAs to 0
# Median imputation is used (rather than mean) because it's robust to outliers in
# financial variables like volume and drawdown.

build_model_frame <- function(train_data, apply_data, feature_cols, binary_features = character(), standardize = FALSE) {
  train_frame <- train_data %>%
    select(all_of(feature_cols)) %>%
    mutate(across(everything(), as.numeric))

  apply_frame <- apply_data %>%
    select(all_of(feature_cols)) %>%
    mutate(across(everything(), as.numeric))

  binary_features <- intersect(binary_features, feature_cols)
  continuous_features <- setdiff(feature_cols, binary_features)

  if (length(binary_features) > 0) {
    for (feature_name in binary_features) {
      train_frame[[feature_name]][is.na(train_frame[[feature_name]])] <- 0
      apply_frame[[feature_name]][is.na(apply_frame[[feature_name]])] <- 0
    }
  }

  if (length(continuous_features) > 0) {
    for (feature_name in continuous_features) {
      median_value <- median(train_frame[[feature_name]], na.rm = TRUE)
      if (!is.finite(median_value)) {
        median_value <- 0
      }

      train_frame[[feature_name]][is.na(train_frame[[feature_name]])] <- median_value
      apply_frame[[feature_name]][is.na(apply_frame[[feature_name]])] <- median_value

      if (standardize) {
        mean_value <- mean(train_frame[[feature_name]], na.rm = TRUE)
        sd_value <- sd(train_frame[[feature_name]], na.rm = TRUE)
        if (!is.finite(sd_value) || sd_value == 0) {
          sd_value <- 1
        }

        train_frame[[feature_name]] <- (train_frame[[feature_name]] - mean_value) / sd_value
        apply_frame[[feature_name]] <- (apply_frame[[feature_name]] - mean_value) / sd_value
      }
    }
  }

  list(
    train = train_frame,
    apply = apply_frame
  )
}


```

## Machine Learning Risk Warning — Two-Year External Validation

The two-year warning extends the one-year setup with a longer forecast window:
a firm is flagged if it crashes in **either** of the next two years. This
reduces false negatives at the cost of more false positives, which may be
acceptable for screening systems that prioritize recall over precision.

The pipeline mirrors the one-year structure but with:
- Two-year labels: `future_2y_crash_dummy = crash_t1 | crash_t2`
- Two-year price extensions (2024 + 2025 fetched)
- Feature-year CV instead of target-year CV
- Confusion matrix outputs for all models

### Two-Year Machine Learning Helper Functions
```{r ml-warning-2y-helpers}
# ---- Standardize raw Yahoo Finance CSV into the canonical price format ----
standardize_warning_price_frame <- function(price_raw, price_map) {
  price_raw <- price_raw %>%
    mutate(
      price_symbol = as.character(price_symbol),
      date = as.Date(date),
      adj_close = as.numeric(adj_close),
      volume = as.numeric(volume)
    ) %>%
    filter(!is.na(adj_close))

  price_has_metadata <- all(
    c("symbol", "company_name", "cik", "gics_sector", "gics_sub_industry") %in% names(price_raw)
  )

  if (price_has_metadata) {
    price_raw %>%
      transmute(
        price_symbol = price_symbol,
        date = date,
        adj_close = adj_close,
        volume = volume,
        symbol = as.character(symbol),
        company_name = as.character(company_name),
        cik = cik,
        gics_sector = as.character(gics_sector),
        gics_sub_industry = as.character(gics_sub_industry)
      ) %>%
      arrange(symbol, date)
  } else {
    price_raw %>%
      select(price_symbol, date, adj_close, volume) %>%
      inner_join(price_map, by = c("price_symbol" = "symbol")) %>%
      transmute(
        price_symbol = price_symbol,
        date = date,
        adj_close = adj_close,
        volume = volume,
        symbol = price_symbol,
        company_name = as.character(company_name),
        cik = cik,
        gics_sector = as.character(gics_sector),
        gics_sub_industry = as.character(gics_sub_industry)
      ) %>%
      arrange(symbol, date)
  }
}

# ---- Fetch or load a single year of daily prices, caching to CSV ----
ensure_warning_price_year <- function(year_value, output_file, price_map, force_refresh = FALSE) {
  start_date <- sprintf("%d-01-01", year_value)
  end_date_exclusive <- sprintf("%d-01-01", year_value + 1L)

  if (!file.exists(output_file) || force_refresh) {
    message(sprintf(">>> Fetching %d daily prices through the Python Yahoo Finance fallback", year_value))
    fetch_log <- fetch_yahoo_prices_python(
      tickers = price_map$symbol,
      start_date = start_date,
      end_date_exclusive = end_date_exclusive,
      output_file = output_file
    )
    message(paste(fetch_log, collapse = "\n"))
  }

  standardized_prices <- readr::read_csv(output_file, show_col_types = FALSE) %>%
    standardize_warning_price_frame(price_map = price_map)

  data.table::fwrite(standardized_prices, output_file)
  standardized_prices
}

calc_classification_metrics <- function(actual, probability, threshold = 0.5) {
  valid_index <- !is.na(actual) & !is.na(probability)
  actual <- as.integer(actual[valid_index])
  probability <- as.numeric(probability[valid_index])

  if (length(actual) == 0) {
    return(tibble())
  }

  predicted_class <- ifelse(probability >= threshold, 1L, 0L)
  true_positive <- sum(predicted_class == 1 & actual == 1)
  true_negative <- sum(predicted_class == 0 & actual == 0)
  false_positive <- sum(predicted_class == 1 & actual == 0)
  false_negative <- sum(predicted_class == 0 & actual == 1)

  precision <- if ((true_positive + false_positive) == 0) NA_real_ else true_positive / (true_positive + false_positive)
  recall <- if ((true_positive + false_negative) == 0) NA_real_ else true_positive / (true_positive + false_negative)
  specificity <- if ((true_negative + false_positive) == 0) NA_real_ else true_negative / (true_negative + false_positive)
  balanced_accuracy <- if (all(is.na(c(recall, specificity)))) {
    NA_real_
  } else {
    mean(c(recall, specificity), na.rm = TRUE)
  }
  top20_summary <- calc_top_n_hit_summary(actual, probability, top_n = 20)

  tibble(
    selected_threshold = threshold,
    n_eval = length(actual),
    positive_count = sum(actual == 1),
    positive_rate = mean(actual == 1),
    predicted_positive_count = sum(predicted_class == 1),
    accuracy = (true_positive + true_negative) / length(actual),
    balanced_accuracy = balanced_accuracy,
    precision = precision,
    recall = recall,
    specificity = specificity,
    f0_5 = calc_fbeta(precision, recall, beta_weight = 0.5),
    roc_auc = calc_roc_auc(actual, probability),
    pr_auc = calc_pr_auc(actual, probability),
    precision_at_top10 = calc_precision_at_top_share(actual, probability, share = 0.10),
    top20_hit_rate = top20_summary$rate,
    hits_in_top20 = top20_summary$hits,
    top20_evaluated_n = top20_summary$k,
    true_positive = true_positive,
    true_negative = true_negative,
    false_positive = false_positive,
    false_negative = false_negative
  )
}

build_threshold_metrics <- function(actual, probability, thresholds) {
  bind_rows(lapply(thresholds, function(one_threshold) {
    calc_classification_metrics(actual, probability, threshold = one_threshold) %>%
      mutate(threshold = one_threshold)
  }))
}

# ---- Build a labeled confusion matrix for model comparison ----
build_confusion_matrix_table <- function(actual, probability, threshold = 0.5) {
  valid_index <- !is.na(actual) & !is.na(probability)
  actual <- as.integer(actual[valid_index])
  probability <- as.numeric(probability[valid_index])

  if (length(actual) == 0) {
    return(tibble())
  }

  predicted_class <- ifelse(probability >= threshold, 1L, 0L)
  true_positive <- sum(predicted_class == 1 & actual == 1)
  true_negative <- sum(predicted_class == 0 & actual == 0)
  false_positive <- sum(predicted_class == 1 & actual == 0)
  false_negative <- sum(predicted_class == 0 & actual == 1)

  tibble(
    selected_threshold = threshold,
    actual_class = c("Actual Crash", "Actual No Crash", "Actual No Crash", "Actual Crash"),
    predicted_class = c("Predicted Crash", "Predicted No Crash", "Predicted Crash", "Predicted No Crash"),
    cell = c("TP", "TN", "FP", "FN"),
    count = c(true_positive, true_negative, false_positive, false_negative)
  ) %>%
    mutate(
      n_eval = sum(count),
      share = ifelse(n_eval > 0, count / n_eval, NA_real_)
    )
}

fit_weighted_logistic_model_target <- function(train_data, apply_data, feature_cols, target_col, binary_features = character()) {
  prepared_frames <- build_model_frame(
    train_data = train_data,
    apply_data = apply_data,
    feature_cols = feature_cols,
    binary_features = binary_features,
    standardize = TRUE
  )

  target_vector <- as.integer(train_data[[target_col]])
  positive_count <- sum(target_vector == 1, na.rm = TRUE)
  negative_count <- sum(target_vector == 0, na.rm = TRUE)

  if (positive_count == 0 || negative_count == 0) {
    return(NULL)
  }

  case_weights <- ifelse(target_vector == 1, negative_count / positive_count, 1)
  fit_data <- bind_cols(tibble(target = target_vector), prepared_frames$train)

  glm_fit <- tryCatch(
    glm(target ~ ., data = fit_data, family = binomial(), weights = case_weights),
    error = function(error) NULL
  )

  if (is.null(glm_fit)) {
    return(NULL)
  }

  predicted_probability <- as.numeric(predict(glm_fit, newdata = prepared_frames$apply, type = "response"))
  predicted_probability <- pmin(pmax(predicted_probability, 1e-6), 1 - 1e-6)

  coefficient_table <- summary(glm_fit)$coefficients

  list(
    probability = predicted_probability,
    fitted_model = glm_fit,
    coefficient_table = tibble(
      term = rownames(coefficient_table),
      estimate = as.numeric(coefficient_table[, 1]),
      std_error = as.numeric(coefficient_table[, 2]),
      statistic = as.numeric(coefficient_table[, 3]),
      p_value = as.numeric(coefficient_table[, 4])
    )
  )
}

fit_ranger_probability_model_target <- function(train_data, apply_data, feature_cols, target_col, binary_features = character()) {
  prepared_frames <- build_model_frame(
    train_data = train_data,
    apply_data = apply_data,
    feature_cols = feature_cols,
    binary_features = binary_features,
    standardize = FALSE
  )

  target_vector <- as.integer(train_data[[target_col]])
  positive_count <- sum(target_vector == 1, na.rm = TRUE)
  negative_count <- sum(target_vector == 0, na.rm = TRUE)

  if (positive_count == 0 || negative_count == 0) {
    return(NULL)
  }

  case_weights <- ifelse(target_vector == 1, negative_count / positive_count, 1)
  fit_data <- bind_cols(
    tibble(target = factor(target_vector, levels = c(0, 1))),
    prepared_frames$train
  )

  ranger_fit <- tryCatch(
    ranger::ranger(
      target ~ .,
      data = fit_data,
      probability = TRUE,
      num.trees = 800,
      importance = "impurity",
      case.weights = case_weights,
      seed = 20260425,
      mtry = max(1, floor(sqrt(length(feature_cols))))
    ),
    error = function(error) NULL
  )

  if (is.null(ranger_fit)) {
    return(NULL)
  }

  prediction_matrix <- predict(ranger_fit, data = prepared_frames$apply)$predictions
  probability_column <- if ("1" %in% colnames(prediction_matrix)) prediction_matrix[, "1"] else prediction_matrix[, ncol(prediction_matrix)]

  list(
    probability = as.numeric(probability_column),
    fitted_model = ranger_fit,
    importance_table = tibble(
      term = names(ranger_fit$variable.importance),
      importance = as.numeric(ranger_fit$variable.importance)
    ) %>%
      arrange(desc(importance))
  )
}
```

### Build 2024-2025 Prices and Two-Year Labels

```{r ml-warning-2y-prices-labels}
price_map_ml <- final_panel %>%
  distinct(symbol, company_name, cik, gics_sector, gics_sub_industry)

price_2024 <- ensure_warning_price_year(
  year_value = WARNING_EXTERNAL_START_YEAR,
  output_file = DERIVED_ML_FILES$price_2024,
  price_map = price_map_ml,
  force_refresh = FORCE_REFRESH_WARNING_PRICES
)

price_2025 <- ensure_warning_price_year(
  year_value = WARNING_EXTERNAL_END_YEAR,
  output_file = DERIVED_ML_FILES$price_2025,
  price_map = price_map_ml,
  force_refresh = FORCE_REFRESH_WARNING_PRICES
)

price_daily_extended <- bind_rows(
  price_daily %>%
    select(price_symbol, date, adj_close, volume, symbol, company_name, cik, gics_sector, gics_sub_industry),
  price_2024,
  price_2025
) %>%
  distinct(symbol, date, .keep_all = TRUE) %>%
  arrange(symbol, date)

data.table::fwrite(price_daily_extended, DERIVED_ML_FILES$price_daily_extended)

if (!file.exists(DERIVED_ML_FILES$crash_metrics_extended) || FORCE_REBUILD_WARNING_LABELS) {
  message(">>> Rebuilding crash metrics through 2025")
  crash_metrics_extended <- build_sp500_crash_metrics(price_daily_extended)
  data.table::fwrite(crash_metrics_extended, DERIVED_ML_FILES$crash_metrics_extended)
} else {
  crash_metrics_extended <- readr::read_csv(
    DERIVED_ML_FILES$crash_metrics_extended,
    show_col_types = FALSE
  )
}

crash_metrics_extended <- crash_metrics_extended %>%
  mutate(
    year = as.integer(year),
    NCSKEW = as.numeric(NCSKEW),
    CRASH_DUMMY = as.integer(CRASH_DUMMY),
    SIGMA = as.numeric(SIGMA),
    RET_MEAN = as.numeric(RET_MEAN),
    DTURN = as.numeric(DTURN)
  ) %>%
  arrange(symbol, year)

two_year_label_summary <- crash_metrics_extended %>%
  filter(year >= 2013, year <= WARNING_EXTERNAL_END_YEAR) %>%
  group_by(year) %>%
  summarise(
    firms_with_label = n(),
    crash_rate = mean(CRASH_DUMMY, na.rm = TRUE),
    mean_ncskew = mean(NCSKEW, na.rm = TRUE),
    .groups = "drop"
  )

two_year_label_summary %>% knitr::kable(digits = 4)
```

### Build the Two-Year Supervised Learning Sample

The two-year sample pairs feature year (t) with a two-year-ahead window
[t+1, t+2]. A firm is labeled positive if it experiences a crash in either year.
The label construction:
  - `crash_t1`: crash event in t+1
  - `crash_t2`: crash event in t+2
  - `future_2y_crash_dummy`: crash_t1 | crash_t2 (only when both years are observed)

```{r ml-warning-2y-sample}
# ---- Reuse the same feature column groups from the one-year setup ----
price_feature_cols <- c(
  "annual_return",
  "annual_log_return",
  "annual_volatility",
  "max_drawdown",
  "avg_volume",
  "median_volume",
  "n_trading_days"
)

snapshot_numeric_cols <- c(
  "snapshot_total_esg_risk",
  "snapshot_environment_risk",
  "snapshot_social_risk",
  "snapshot_governance_risk",
  "snapshot_controversy_score",
  "snapshot_esg_percentile"
)

edgar_numeric_cols <- c(
  "edgar_e_score",
  "edgar_s_score",
  "edgar_g_score",
  "edgar_esg_total",
  "edgar_word_count"
)

warning_feature_panel_2y <- panel_for_usability %>%
  transmute(
    symbol,
    company_name,
    sector_standardized,
    feature_year = year,
    window_start_year = year + 1L,
    window_end_year = year + 2L,
    eligible_by_year,
    across(all_of(price_feature_cols)),
    snapshot_available = as.integer(snapshot_available),
    edgar_available = as.integer(edgar_available),
    across(all_of(snapshot_numeric_cols)),
    across(all_of(edgar_numeric_cols))
  )

crash_label_t1 <- crash_metrics_extended %>%
  transmute(
    symbol,
    window_start_year = year,
    crash_t1 = as.integer(CRASH_DUMMY),
    ncskew_t1 = as.numeric(NCSKEW)
  )

crash_label_t2 <- crash_metrics_extended %>%
  transmute(
    symbol,
    window_end_year = year,
    crash_t2 = as.integer(CRASH_DUMMY),
    ncskew_t2 = as.numeric(NCSKEW)
  )

# ---- Merge t+1 and t+2 crash labels into a single two-year outcome ----
ml_warning_panel_2y <- warning_feature_panel_2y %>%
  left_join(crash_label_t1, by = c("symbol", "window_start_year")) %>%
  left_join(crash_label_t2, by = c("symbol", "window_end_year")) %>%
  mutate(
    across(all_of(snapshot_numeric_cols), ~ suppressWarnings(readr::parse_number(as.character(.x)))),
    across(all_of(edgar_numeric_cols), ~ suppressWarnings(readr::parse_number(as.character(.x)))),
    across(all_of(price_feature_cols), ~ suppressWarnings(readr::parse_number(as.character(.x)))),
    ncskew_t1 = suppressWarnings(readr::parse_number(as.character(ncskew_t1))),
    ncskew_t2 = suppressWarnings(readr::parse_number(as.character(ncskew_t2))),
    crash_t1 = as.integer(crash_t1),
    crash_t2 = as.integer(crash_t2)
  )

for (feature_name in snapshot_numeric_cols) {
  ml_warning_panel_2y[[paste0(feature_name, "_is_missing")]] <- as.integer(is.na(ml_warning_panel_2y[[feature_name]]))
}

for (feature_name in edgar_numeric_cols) {
  ml_warning_panel_2y[[paste0(feature_name, "_is_missing")]] <- as.integer(is.na(ml_warning_panel_2y[[feature_name]]))
}

ml_warning_panel_2y <- ml_warning_panel_2y %>%
  mutate(
    window_complete = !is.na(crash_t1) & !is.na(crash_t2),
    future_2y_crash_dummy = if_else(
      window_complete,
      as.integer(crash_t1 == 1L | crash_t2 == 1L),
      NA_integer_
    ),
    data_split = case_when(
      feature_year >= START_YEAR & feature_year <= (END_YEAR - 1L) ~ "pre_external_training_pool",
      feature_year == END_YEAR ~ "external_2024_2025_test",
      TRUE ~ "unused"
    )
  ) %>%
  filter(feature_year >= START_YEAR, window_end_year <= WARNING_EXTERNAL_END_YEAR)

training_pool <- ml_warning_panel_2y %>%
  filter(
    eligible_by_year,
    window_complete,
    data_split == "pre_external_training_pool"
  )

external_test_pool <- ml_warning_panel_2y %>%
  filter(
    eligible_by_year,
    window_complete,
    feature_year == END_YEAR,
    data_split == "external_2024_2025_test"
  )

data.table::fwrite(ml_warning_panel_2y, DERIVED_ML_FILES$ml_panel)
data.table::fwrite(training_pool, OUTPUT_ML_FILES$training_panel)

tibble(
  sample_name = c("training_pool_2y", "external_test_pool_2y"),
  rows = c(nrow(training_pool), nrow(external_test_pool)),
  companies = c(dplyr::n_distinct(training_pool$symbol), dplyr::n_distinct(external_test_pool$symbol)),
  positive_rate = c(mean(training_pool$future_2y_crash_dummy, na.rm = TRUE), mean(external_test_pool$future_2y_crash_dummy, na.rm = TRUE)),
  feature_year_span = c(
    paste(range(training_pool$feature_year, na.rm = TRUE), collapse = "-"),
    paste(range(external_test_pool$feature_year, na.rm = TRUE), collapse = "-")
  ),
  window_span = c(
    paste0(min(training_pool$window_start_year, na.rm = TRUE), "-", max(training_pool$window_end_year, na.rm = TRUE)),
    paste0(min(external_test_pool$window_start_year, na.rm = TRUE), "-", max(external_test_pool$window_end_year, na.rm = TRUE))
  )
) %>%
  knitr::kable(digits = 4)
```

### Two-Year Model Training and Threshold Selection

Training uses **feature-year CV** (2018-2022): models train on earlier feature years
and validate on each holdout feature year. This differs from the one-year setup
(target-year CV) because each feature year's validation window spans two future years,
making target-year splits less clean.

The same 6 model specifications (3 feature sets × 2 algorithms) are reused so
one-year and two-year results are directly comparable.

```{r ml-warning-2y-training}
snapshot_missing_flags <- paste0(snapshot_numeric_cols, "_is_missing")
edgar_missing_flags <- paste0(edgar_numeric_cols, "_is_missing")

# ---- Mirror the one-year specifications (same feature sets, same algorithms) ----
model_specifications <- tibble(
  feature_set_id = c("price_baseline", "price_plus_snapshot_esg", "price_plus_full_esg"),
  feature_set_label = c("Price Baseline", "Price + Snapshot ESG", "Price + Full ESG"),
  feature_columns = list(
    price_feature_cols,
    c(price_feature_cols, "snapshot_available", snapshot_numeric_cols, snapshot_missing_flags),
    c(price_feature_cols, "snapshot_available", snapshot_numeric_cols, snapshot_missing_flags, "edgar_available", edgar_numeric_cols, edgar_missing_flags)
  ),
  binary_columns = list(
    character(),
    c("snapshot_available", snapshot_missing_flags),
    c("snapshot_available", snapshot_missing_flags, "edgar_available", edgar_missing_flags)
  )
) %>%
  tidyr::crossing(
    tibble(
      model_type = c("weighted_logistic", "ranger_probability_forest"),
      model_label = c("Weighted Logistic", "Probability Random Forest")
    )
  ) %>%
  mutate(
    model_key = paste(feature_set_id, model_type, sep = "__"),
    model_display = paste(feature_set_label, model_label, sep = " | ")
  )

# ---- Feature-year CV: validate on 2018-2022 features, reserve 2023 for external test ----
validation_feature_years <- 2018:2022
threshold_grid <- seq(0.05, 0.95, by = 0.01)

validation_fold_results <- list()
validation_threshold_results <- list()

for (spec_index in seq_len(nrow(model_specifications))) {
  one_spec <- model_specifications[spec_index, ]
  feature_columns <- one_spec$feature_columns[[1]]
  binary_columns <- one_spec$binary_columns[[1]]

  for (one_validation_feature_year in validation_feature_years) {
    train_fold <- training_pool %>%
      filter(feature_year < one_validation_feature_year)

    valid_fold <- training_pool %>%
      filter(feature_year == one_validation_feature_year)

    if (nrow(train_fold) == 0 || nrow(valid_fold) == 0) {
      next
    }

    fitted_output <- if (one_spec$model_type[[1]] == "weighted_logistic") {
      fit_weighted_logistic_model_target(
        train_data = train_fold,
        apply_data = valid_fold,
        feature_cols = feature_columns,
        target_col = "future_2y_crash_dummy",
        binary_features = binary_columns
      )
    } else {
      fit_ranger_probability_model_target(
        train_data = train_fold,
        apply_data = valid_fold,
        feature_cols = feature_columns,
        target_col = "future_2y_crash_dummy",
        binary_features = binary_columns
      )
    }

    if (is.null(fitted_output)) {
      next
    }

    fold_prediction <- valid_fold %>%
      transmute(
        model_key = one_spec$model_key[[1]],
        model_display = one_spec$model_display[[1]],
        feature_set_id = one_spec$feature_set_id[[1]],
        feature_set_label = one_spec$feature_set_label[[1]],
        model_type = one_spec$model_type[[1]],
        model_label = one_spec$model_label[[1]],
        validation_feature_year = one_validation_feature_year,
        validation_window = paste0(window_start_year, "-", window_end_year),
        symbol,
        company_name,
        actual_label = future_2y_crash_dummy,
        crash_t1,
        crash_t2,
        risk_probability = fitted_output$probability
      ) %>%
      mutate(
        risk_rank = min_rank(desc(risk_probability))
      )

    fold_metrics <- calc_classification_metrics(
      actual = fold_prediction$actual_label,
      probability = fold_prediction$risk_probability,
      threshold = 0.50
    ) %>%
      mutate(
        model_key = one_spec$model_key[[1]],
        model_display = one_spec$model_display[[1]],
        feature_set_id = one_spec$feature_set_id[[1]],
        feature_set_label = one_spec$feature_set_label[[1]],
        model_type = one_spec$model_type[[1]],
        model_label = one_spec$model_label[[1]],
        validation_feature_year = one_validation_feature_year,
        validation_window = paste0(min(valid_fold$window_start_year), "-", max(valid_fold$window_end_year)),
        train_rows = nrow(train_fold),
        validation_rows = nrow(valid_fold),
        train_positive_rate = mean(train_fold$future_2y_crash_dummy, na.rm = TRUE),
        validation_positive_rate = mean(valid_fold$future_2y_crash_dummy, na.rm = TRUE)
      )

    threshold_metrics <- build_threshold_metrics(
      actual = fold_prediction$actual_label,
      probability = fold_prediction$risk_probability,
      thresholds = threshold_grid
    ) %>%
      mutate(
        model_key = one_spec$model_key[[1]],
        model_display = one_spec$model_display[[1]],
        feature_set_id = one_spec$feature_set_id[[1]],
        feature_set_label = one_spec$feature_set_label[[1]],
        model_type = one_spec$model_type[[1]],
        model_label = one_spec$model_label[[1]],
        validation_feature_year = one_validation_feature_year
      )

    validation_fold_results[[length(validation_fold_results) + 1]] <- fold_metrics
    validation_threshold_results[[length(validation_threshold_results) + 1]] <- threshold_metrics
  }
}

validation_fold_metrics <- bind_rows(validation_fold_results)
validation_threshold_metrics <- bind_rows(validation_threshold_results)

threshold_selection <- validation_threshold_metrics %>%
  group_by(model_key, model_display, feature_set_id, feature_set_label, model_type, model_label, threshold) %>%
  summarise(
    mean_accuracy = mean(accuracy, na.rm = TRUE),
    mean_balanced_accuracy = mean(balanced_accuracy, na.rm = TRUE),
    mean_precision = mean(precision, na.rm = TRUE),
    mean_recall = mean(recall, na.rm = TRUE),
    mean_f0_5 = mean(f0_5, na.rm = TRUE),
    mean_pr_auc_proxy = mean(pr_auc, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(model_key, model_display, feature_set_id, feature_set_label, model_type, model_label) %>%
  group_modify(~ choose_threshold_from_validation(.x)) %>%
  ungroup() %>%
  rename(selected_threshold = threshold)

cv_model_comparison <- validation_fold_metrics %>%
  group_by(model_key, model_display, feature_set_id, feature_set_label, model_type, model_label) %>%
  summarise(
    fold_count = n(),
    mean_accuracy = mean(accuracy, na.rm = TRUE),
    mean_balanced_accuracy = mean(balanced_accuracy, na.rm = TRUE),
    mean_precision = mean(precision, na.rm = TRUE),
    mean_recall = mean(recall, na.rm = TRUE),
    mean_f0_5 = mean(f0_5, na.rm = TRUE),
    mean_roc_auc = mean(roc_auc, na.rm = TRUE),
    mean_pr_auc = mean(pr_auc, na.rm = TRUE),
    mean_precision_at_top10 = mean(precision_at_top10, na.rm = TRUE),
    mean_top20_hit_rate = mean(top20_hit_rate, na.rm = TRUE),
    mean_train_positive_rate = mean(train_positive_rate, na.rm = TRUE),
    mean_validation_positive_rate = mean(validation_positive_rate, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(
    threshold_selection %>%
      select(model_key, selected_threshold, mean_accuracy, mean_balanced_accuracy, mean_precision, mean_recall, mean_f0_5),
    by = "model_key",
    suffix = c("_cv", "_threshold")
  ) %>%
  arrange(desc(mean_pr_auc), desc(mean_precision_at_top10), desc(mean_f0_5_cv), desc(mean_accuracy_cv), model_type)

best_overall_model <- cv_model_comparison %>%
  arrange(desc(mean_pr_auc), desc(mean_precision_at_top10), desc(mean_f0_5_cv), desc(mean_accuracy_cv), model_type) %>%
  slice(1)

best_logistic_model <- cv_model_comparison %>%
  filter(model_type == "weighted_logistic") %>%
  arrange(desc(mean_pr_auc), desc(mean_precision_at_top10), desc(mean_f0_5_cv), desc(mean_accuracy_cv)) %>%
  slice(1)

best_forest_model <- cv_model_comparison %>%
  filter(model_type == "ranger_probability_forest") %>%
  arrange(desc(mean_pr_auc), desc(mean_precision_at_top10), desc(mean_f0_5_cv), desc(mean_accuracy_cv)) %>%
  slice(1)

data.table::fwrite(validation_fold_metrics, OUTPUT_ML_FILES$validation_fold_metrics)
data.table::fwrite(validation_threshold_metrics, OUTPUT_ML_FILES$threshold_tuning)
data.table::fwrite(cv_model_comparison, OUTPUT_ML_FILES$cv_model_comparison)

cv_model_comparison %>%
  select(model_display, mean_pr_auc, mean_precision_at_top10, mean_f0_5_cv, mean_accuracy_cv, selected_threshold) %>%
  knitr::kable(digits = 4)
```

### 2024-2025 Two-Year External Validation and Warning Outputs

The two-year external test uses 2023 features to predict crashes in the 2024-2025
window. Since 2025 crash labels are now realized, this provides an additional
out-of-sample check beyond the one-year 2024-only test.

Additional outputs vs. one-year:
- Confusion matrices (all models + champion only)
- Accuracy/precision/recall/F0.5 comparison bar chart
- Two-year champion model's top-20 high-risk company list

```{r ml-warning-2y-external-validation}
external_prediction_results <- list()
logistic_coefficient_tables <- list()
forest_importance_tables <- list()
pr_curve_tables <- list()

for (spec_index in seq_len(nrow(model_specifications))) {
  one_spec <- model_specifications[spec_index, ]
  feature_columns <- one_spec$feature_columns[[1]]
  binary_columns <- one_spec$binary_columns[[1]]

  fitted_output <- if (one_spec$model_type[[1]] == "weighted_logistic") {
    fit_weighted_logistic_model_target(
      train_data = training_pool,
      apply_data = external_test_pool,
      feature_cols = feature_columns,
      target_col = "future_2y_crash_dummy",
      binary_features = binary_columns
    )
  } else {
    fit_ranger_probability_model_target(
      train_data = training_pool,
      apply_data = external_test_pool,
      feature_cols = feature_columns,
      target_col = "future_2y_crash_dummy",
      binary_features = binary_columns
    )
  }

  if (is.null(fitted_output)) {
    next
  }

  final_threshold <- threshold_selection %>%
    filter(model_key == one_spec$model_key[[1]]) %>%
    pull(selected_threshold)

  if (length(final_threshold) == 0 || is.na(final_threshold[[1]])) {
    final_threshold <- 0.5
  } else {
    final_threshold <- final_threshold[[1]]
  }

  final_prediction <- external_test_pool %>%
    transmute(
      model_key = one_spec$model_key[[1]],
      model_display = one_spec$model_display[[1]],
      feature_set_id = one_spec$feature_set_id[[1]],
      feature_set_label = one_spec$feature_set_label[[1]],
      model_type = one_spec$model_type[[1]],
      model_label = one_spec$model_label[[1]],
      symbol,
      company_name,
      sector_standardized,
      feature_year,
      window_start_year,
      window_end_year,
      crash_t1,
      crash_t2,
      ncskew_t1,
      ncskew_t2,
      future_2y_crash_dummy,
      risk_probability = fitted_output$probability
    ) %>%
    mutate(
      selected_threshold = final_threshold,
      warning_flag = risk_probability >= selected_threshold,
      risk_rank = min_rank(desc(risk_probability))
    )

  external_prediction_results[[length(external_prediction_results) + 1]] <- final_prediction

  if (!is.null(fitted_output$coefficient_table)) {
    logistic_coefficient_tables[[one_spec$model_key[[1]]]] <- fitted_output$coefficient_table %>%
      mutate(
        model_key = one_spec$model_key[[1]],
        model_display = one_spec$model_display[[1]]
      )
  }

  if (!is.null(fitted_output$importance_table)) {
    forest_importance_tables[[one_spec$model_key[[1]]]] <- fitted_output$importance_table %>%
      mutate(
        model_key = one_spec$model_key[[1]],
        model_display = one_spec$model_display[[1]]
      )
  }

  pr_curve_tables[[one_spec$model_key[[1]]]] <- build_pr_curve_data(
    actual = final_prediction$future_2y_crash_dummy,
    probability = final_prediction$risk_probability,
    model_label = one_spec$model_display[[1]]
  )
}

external_predictions <- bind_rows(external_prediction_results)
logistic_coefficients_all <- bind_rows(logistic_coefficient_tables)
forest_importance_all <- bind_rows(forest_importance_tables)
pr_curve_2y <- bind_rows(pr_curve_tables)

external_test_metrics <- external_predictions %>%
  group_by(model_key, model_display, feature_set_id, feature_set_label, model_type, model_label) %>%
  group_modify(~ {
    threshold_value <- unique(.x$selected_threshold)[1]
    calc_classification_metrics(.x$future_2y_crash_dummy, .x$risk_probability, threshold = threshold_value)
  }) %>%
  ungroup() %>%
  arrange(desc(pr_auc), desc(precision_at_top10), desc(f0_5), desc(accuracy))

confusion_matrix_by_model <- external_predictions %>%
  group_by(model_key, model_display, feature_set_id, feature_set_label, model_type, model_label) %>%
  group_modify(~ {
    threshold_value <- unique(.x$selected_threshold)[1]
    build_confusion_matrix_table(.x$future_2y_crash_dummy, .x$risk_probability, threshold = threshold_value)
  }) %>%
  ungroup()

champion_model_key <- best_overall_model$model_key[[1]]
champion_predictions <- external_predictions %>%
  filter(model_key == champion_model_key) %>%
  arrange(risk_rank)

champion_model_label <- if (nrow(champion_predictions) > 0) champion_predictions$model_display[[1]] else "No model available"

top_risk_companies_2y <- champion_predictions %>%
  slice_head(n = 20) %>%
  select(
    model_display,
    symbol,
    company_name,
    sector_standardized,
    feature_year,
    window_start_year,
    window_end_year,
    risk_probability,
    selected_threshold,
    warning_flag,
    crash_t1,
    crash_t2,
    future_2y_crash_dummy,
    risk_rank
  )

baseline_external <- external_test_metrics %>%
  filter(feature_set_id == "price_baseline") %>%
  arrange(desc(pr_auc), desc(precision_at_top10), desc(f0_5), desc(accuracy)) %>%
  slice(1)

best_esg_external <- external_test_metrics %>%
  filter(feature_set_id != "price_baseline") %>%
  arrange(desc(pr_auc), desc(precision_at_top10), desc(f0_5), desc(accuracy)) %>%
  slice(1)

esg_gain_summary <- tibble(
  baseline_model = baseline_external$model_display,
  baseline_pr_auc = baseline_external$pr_auc,
  baseline_precision_at_top10 = baseline_external$precision_at_top10,
  best_esg_model = best_esg_external$model_display,
  best_esg_pr_auc = best_esg_external$pr_auc,
  best_esg_precision_at_top10 = best_esg_external$precision_at_top10,
  esg_adds_value = ifelse(
    best_esg_external$pr_auc > baseline_external$pr_auc |
      best_esg_external$precision_at_top10 > baseline_external$precision_at_top10,
    "YES",
    "NO"
  )
)

data.table::fwrite(external_test_metrics, OUTPUT_ML_FILES$test_metrics)
data.table::fwrite(confusion_matrix_by_model, OUTPUT_ML_FILES$confusion_matrix)
data.table::fwrite(external_predictions, OUTPUT_ML_FILES$company_predictions)
data.table::fwrite(top_risk_companies_2y, OUTPUT_ML_FILES$top_risk_companies)

pr_curve_plot <- pr_curve_2y %>%
  ggplot(aes(x = recall, y = precision, color = model_label)) +
  geom_line(linewidth = 1) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "2024-2025 External Test PR Curve",
    x = "Recall",
    y = "Precision",
    color = "Model"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")

accuracy_precision_plot_data <- external_test_metrics %>%
  select(model_display, accuracy, balanced_accuracy, precision, recall, f0_5) %>%
  pivot_longer(
    cols = -model_display,
    names_to = "metric",
    values_to = "value"
  ) %>%
  mutate(
    metric = factor(
      metric,
      levels = c("accuracy", "balanced_accuracy", "precision", "recall", "f0_5"),
      labels = c("Accuracy", "Balanced Accuracy", "Precision", "Recall", "F0.5")
    ),
    model_display = factor(model_display, levels = rev(unique(external_test_metrics$model_display)))
  )

accuracy_precision_plot <- accuracy_precision_plot_data %>%
  ggplot(aes(x = model_display, y = value, fill = metric)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.72) +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "2024-2025 External Test Accuracy and Precision Profile",
    x = "Model",
    y = "Metric Value",
    fill = "Metric"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")

confusion_matrix_plot_data <- confusion_matrix_by_model %>%
  mutate(
    actual_class = factor(actual_class, levels = c("Actual Crash", "Actual No Crash")),
    predicted_class = factor(predicted_class, levels = c("Predicted Crash", "Predicted No Crash")),
    cell_label = paste0("n = ", count, "\n", scales::percent(share, accuracy = 0.1))
  )

confusion_matrix_all_plot <- confusion_matrix_plot_data %>%
  ggplot(aes(x = predicted_class, y = actual_class, fill = share)) +
  geom_tile(color = "white", linewidth = 0.6) +
  geom_text(aes(label = cell_label), size = 3.2, lineheight = 0.95) +
  facet_wrap(~ model_display) +
  scale_fill_gradient(low = "#F3F4F6", high = "#2F6DB3", labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "2024-2025 Confusion Matrix by Model",
    x = "Predicted Class",
    y = "Actual Class",
    fill = "Share"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid = element_blank()
  )

champion_confusion_plot <- confusion_matrix_plot_data %>%
  filter(model_key == champion_model_key) %>%
  ggplot(aes(x = predicted_class, y = actual_class, fill = share)) +
  geom_tile(color = "white", linewidth = 0.8) +
  geom_text(aes(label = cell_label), size = 5, lineheight = 0.95) +
  scale_fill_gradient(low = "#F3F4F6", high = "#2F6DB3", labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Champion Model Confusion Matrix (2024-2025)",
    subtitle = champion_model_label,
    x = "Predicted Class",
    y = "Actual Class",
    fill = "Share"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "bottom",
    panel.grid = element_blank()
  )

top_risk_plot <- top_risk_companies_2y %>%
  mutate(company_name = reorder(company_name, risk_probability)) %>%
  ggplot(aes(x = company_name, y = risk_probability, fill = factor(future_2y_crash_dummy, levels = c(0, 1)))) +
  geom_col(width = 0.75) +
  coord_flip() +
  scale_fill_manual(values = c("0" = "#BDBDBD", "1" = "#D95F02"), labels = c("0" = "No Crash in 2024-2025", "1" = "Crash in 2024-2025")) +
  labs(
    title = "Top 20 High-Risk Companies for 2024-2025",
    subtitle = paste0("Champion model: ", champion_model_label),
    x = "Company",
    y = "Risk Probability",
    fill = "Observed Outcome"
  ) +
  theme_minimal(base_size = 12)

best_logistic_key <- if (nrow(best_logistic_model) > 0) best_logistic_model$model_key[[1]] else NA_character_
best_logistic_label <- if (nrow(best_logistic_model) > 0) best_logistic_model$model_display[[1]] else "No logistic model available"

best_logistic_coefficients <- logistic_coefficients_all %>%
  filter(model_key == best_logistic_key, term != "(Intercept)") %>%
  mutate(abs_estimate = abs(estimate)) %>%
  slice_max(abs_estimate, n = 15, with_ties = FALSE) %>%
  arrange(estimate) %>%
  mutate(term = factor(term, levels = term))

logistic_coef_plot <- if (nrow(best_logistic_coefficients) > 0) {
  best_logistic_coefficients %>%
    ggplot(aes(x = term, y = estimate, fill = estimate > 0)) +
    geom_col(show.legend = FALSE) +
    coord_flip() +
    scale_fill_manual(values = c("TRUE" = "#2F6DB3", "FALSE" = "#D95F02")) +
    labs(
      title = "Best Logistic Model Coefficients (2-Year Warning)",
      subtitle = best_logistic_label,
      x = "Feature",
      y = "Coefficient Estimate"
    ) +
    theme_minimal(base_size = 12)
} else {
  ggplot() +
    annotate("text", x = 1, y = 1, label = "No logistic coefficients available") +
    labs(title = "Best Logistic Model Coefficients (2-Year Warning)") +
    theme_void()
}

best_forest_key <- if (nrow(best_forest_model) > 0) best_forest_model$model_key[[1]] else NA_character_
best_forest_label <- if (nrow(best_forest_model) > 0) best_forest_model$model_display[[1]] else "No forest model available"

best_forest_importance <- forest_importance_all %>%
  filter(model_key == best_forest_key) %>%
  slice_max(importance, n = 15, with_ties = FALSE) %>%
  arrange(importance) %>%
  mutate(term = factor(term, levels = term))

forest_importance_plot <- if (nrow(best_forest_importance) > 0) {
  best_forest_importance %>%
    ggplot(aes(x = term, y = importance)) +
    geom_col(fill = "#2F6DB3", width = 0.75) +
    coord_flip() +
    labs(
      title = "Best Probability Forest Feature Importance (2-Year Warning)",
      subtitle = best_forest_label,
      x = "Feature",
      y = "Importance"
    ) +
    theme_minimal(base_size = 12)
} else {
  ggplot() +
    annotate("text", x = 1, y = 1, label = "No forest importance available") +
    labs(title = "Best Probability Forest Feature Importance (2-Year Warning)") +
    theme_void()
}

ggplot2::ggsave(OUTPUT_ML_FILES$pr_curve, pr_curve_plot, width = 9.5, height = 6.2, dpi = 160)
ggplot2::ggsave(OUTPUT_ML_FILES$accuracy_precision_plot, accuracy_precision_plot, width = 10.5, height = 6.8, dpi = 160)
ggplot2::ggsave(OUTPUT_ML_FILES$confusion_all_plot, confusion_matrix_all_plot, width = 12, height = 8, dpi = 160)
ggplot2::ggsave(OUTPUT_ML_FILES$confusion_champion_plot, champion_confusion_plot, width = 7.5, height = 6.5, dpi = 160)
ggplot2::ggsave(OUTPUT_ML_FILES$top_risk_plot, top_risk_plot, width = 10, height = 7, dpi = 160)
ggplot2::ggsave(OUTPUT_ML_FILES$logistic_coef_plot, logistic_coef_plot, width = 10, height = 7, dpi = 160)
ggplot2::ggsave(OUTPUT_ML_FILES$forest_importance_plot, forest_importance_plot, width = 10, height = 7, dpi = 160)

external_test_metrics %>%
  select(model_display, pr_auc, roc_auc, accuracy, balanced_accuracy, precision, recall, f0_5, precision_at_top10, top20_hit_rate, selected_threshold) %>%
  knitr::kable(digits = 4)

esg_gain_summary %>% knitr::kable(digits = 4)
```

### Two-Year Machine Learning Validation Checks

Assert invariants for the two-year setup:
- Window boundaries align correctly with feature years
- Training pool excludes realized 2025 labels
- 2-year label exactly equals crash_t1 | crash_t2
- Confusion matrix cell counts sum to evaluation sample size

```{r ml-warning-2y-validation}
# ---- Verify two-year temporal alignment ----
stopifnot(all(training_pool$window_start_year == training_pool$feature_year + 1L))
stopifnot(all(training_pool$window_end_year == training_pool$feature_year + WARNING_LOOKAHEAD_YEARS))
stopifnot(!any(training_pool$window_end_year == WARNING_EXTERNAL_END_YEAR))
stopifnot(all(external_test_pool$feature_year == END_YEAR))
stopifnot(all(external_test_pool$window_start_year == WARNING_EXTERNAL_START_YEAR))
stopifnot(all(external_test_pool$window_end_year == WARNING_EXTERNAL_END_YEAR))
stopifnot(!any(is.na(training_pool$future_2y_crash_dummy)))
stopifnot(all(
  ml_warning_panel_2y$future_2y_crash_dummy[ml_warning_panel_2y$window_complete] ==
    as.integer(
      ml_warning_panel_2y$crash_t1[ml_warning_panel_2y$window_complete] == 1L |
        ml_warning_panel_2y$crash_t2[ml_warning_panel_2y$window_complete] == 1L
    )
))
stopifnot(nrow(external_test_metrics) == nrow(model_specifications))
stopifnot(all(external_test_metrics$selected_threshold >= 0.05 & external_test_metrics$selected_threshold <= 0.95))
stopifnot(all(external_test_metrics$accuracy >= 0 & external_test_metrics$accuracy <= 1))
stopifnot(all(external_test_metrics$balanced_accuracy[!is.na(external_test_metrics$balanced_accuracy)] >= 0 & external_test_metrics$balanced_accuracy[!is.na(external_test_metrics$balanced_accuracy)] <= 1))
stopifnot(all(external_test_metrics$pr_auc[!is.na(external_test_metrics$pr_auc)] >= 0 & external_test_metrics$pr_auc[!is.na(external_test_metrics$pr_auc)] <= 1))
stopifnot(all(external_test_metrics$roc_auc[!is.na(external_test_metrics$roc_auc)] >= 0 & external_test_metrics$roc_auc[!is.na(external_test_metrics$roc_auc)] <= 1))
stopifnot(all(external_predictions$risk_rank >= 1))

confusion_validation <- confusion_matrix_by_model %>%
  group_by(model_key) %>%
  summarise(total_count = sum(count), .groups = "drop") %>%
  inner_join(external_test_metrics %>% select(model_key, n_eval), by = "model_key")

stopifnot(all(confusion_validation$total_count == confusion_validation$n_eval))

ml_validation_summary_2y <- tibble(
  check = c(
    "window_start_year always equals feature_year + 1",
    "window_end_year always equals feature_year + 2",
    "Training pool excludes realized 2025 labels",
    "External test uses 2023 features and validates 2024-2025",
    "The 2-year label equals a crash in either of the next two years",
    "Accuracy / Balanced Accuracy / PR AUC / ROC AUC lie within [0, 1]",
    "Confusion matrix cells sum to the external test sample size"
  ),
  result = "PASS"
)

knitr::kable(ml_validation_summary_2y)
```

### Two-Year Machine Learning Output Files

```{r ml-warning-2y-output-files}
ml_output_table_2y <- tibble(
  file = to_project_relative(c(unname(unlist(DERIVED_ML_FILES)), unname(unlist(OUTPUT_ML_FILES)))),
  description = c(
    "Cached 2024 daily price data used for the 2-year warning task",
    "Newly fetched 2025 daily price data",
    "Extended 2011-2025 daily price table",
    "Extended 2012-2025 annual crash label table",
    "ML base panel linking 2012-2023 features to 2-year forward labels",
    "2-year training sample export",
    "2-year validation fold metrics by feature year",
    "2-year threshold search results across validation folds",
    "2-year cross-validation model comparison",
    "2024-2025 external test metrics table",
    "2024-2025 confusion matrix table by model",
    "2024-2025 company-level prediction table",
    "2024-2025 high-risk company list",
    "2024-2025 external test PR curve",
    "2024-2025 accuracy and precision comparison chart",
    "2024-2025 confusion matrix chart for all models",
    "2024-2025 confusion matrix chart for the champion model",
    "2024-2025 Top high-risk companies chart",
    "Best logistic coefficient chart for the 2-year task",
    "Best forest feature importance chart for the 2-year task"
  )
)

knitr::kable(ml_output_table_2y)
```

